# Introduction

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
library(readxl)
library(tidyverse)
library(stringr)
library(plotly)
library(ggpubr)
library(mgcv)
library(RColorBrewer)
library(htmltools)
library(htmlwidgets)
library(gridExtra)
library(grid)
library(DT)
library(ggiraph)
library(cowplot)
library(ggnewscale)
library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

load(paste0(extraction.location,"curated.RData"))
```

```{r data_extraction_from_excelsheets and curation - preliminary work, echo = F, include = F, eval = F }

# extract from excel sheets ----
source("Scripts\\Extract_expert_info.R")
# if all is well this should:
# save ind.matcher ind.matcher.df and expert.data (expert opinion responses from the Delphi forms) 
# as files within the specified extraction.location to be loaded together

# curation ----
source("Scripts\\curation of extracted data.R")
# this one loads the data and does some curation, scaling responses etc
# then saves the dataframes

# then later I loads it (in Italian-American mafiosi accent)

```

```{r respondant colours, echo = F, include = F, eval = T }
respondant_colours <- glasbey.colors(nlevels(respondants_with_something)+1)[-1] %>% as.character() %>% 
    setNames(., respondants_with_something)

```

This document explores expert opinion, gathered from Woodland Trust staff, relating to `r has_completed$indicator_name %>% unique() %>% length()` indicators of woodland ecological condition: their effective measurement, and how those measurements might relate to ecological condition.

> **Woodland ecological condition** is defined as: a wood's potential to promote flourishing wildlife species communities (i.e. ignoring time-lags in wildlife species colonisation).

The indicators were selected based on their perceived correlation with woodland ecological condition, and their feasibility of measurement. The relationship between each indicator and woodland ecological condition was captured by estimating **value functions** ("what good looks like" ) and importance **weightings** for each indicator, based on expert opinion.

<button class="btn btn-primary" type="button" data-bs-toggle="collapse" data-bs-target="#IndicatorsTable" aria-expanded="false" aria-controls="collapseExample">

Toggle Indicators Table

</button>

</p>

::: {#IndicatorsTable .collapse}
::: {.card .card-body}
```{r  indicators table, echo = F }

ind.table <- ind.matcher.df [,1:2]
names(ind.table) <- c("Indicator", "Description")
ind.table %>% 
   datatable(#extensions = c('FixedColumns',"FixedHeader"),
              options = list(scrollX = T, 
                             paging = F,
                             fixedHeader=F,
                             dom = 't',
                             fixedColumns = list(leftColumns = 1) #,
                             # buttons = c('copy', 'csv')
                             ),
              rownames = F#,
              #container = sketch.lcm.areas
              ) %>% 
    formatStyle( columns = c("Indicator"), fontWeight = 'bold', `text-align` = 'left', "white-space" = "nowrap") 

```
:::
:::

Expert opinion on the value function and weight for each indicator was elicited using the Delphi method. The Delphi method is a well-established method designed to collect and distill expert knowledge and has a history of successful use in supporting conservation action where empirical evidence is insufficient and information is needed rapidly to inform decisions. It's transparency and repeatability also add to its appeal.

The Delphi method relies on using repeated surveys where participants are asked the same questions in at least two to three ‘rounds’ . Between rounds participants review and consider the answers given by all panel-members and can then revise their answers, increasing their reliability <!-- (Edwards et al., 2011) --> and identifying areas where agreements and disagreements in understanding occur. <!-- (Buckley, 1995; Kangas, Alho, Kolehmainen, & Mononen, 1998; Landeta, 2006). -->
