# Species richness: native trees & shrubs

## Justification

More tree species typically reflect greater niche space and resilience to pests/disease.

## Measuring

Record the Native tree species species present.

> For the condition score, that list of species should then be compared to a list of appropriate native tree species for that site (ideally pre-defined automatically). This comparison could be a simple proportion (e.g. 70% of site-appropriate species occupied the site), or could potentially be much more complex functions that might consider different weightings for species present based on considerations such as:
>
> -   Rarity and contribution to landscape-scale biodiversity (e.g. [Beta diversity](https://methodsblog.com/2015/05/27/beta_diversity/))
>
> -   Dispersal ability (potentially weighting pioneer species and long-distance dispersers less)
>
> -   Varying degrees of "site appropriateness", however that is defined.

## Challenges {#sec-tree-sr-chalenges}

### Site-appropriateness

A list of site-appropriate species should be defined prior to surveying in a way that is objective and fully repeatable. Ideally this would be automatic. This list could relate to:

-   NVC type or native species predicted at the site by Forest Research's [Ecological Site Classification model (ESC)](https://www.forestresearch.gov.uk/tools-and-resources/fthr/ecological-site-classification/)

-   Historic management, disease and site objectives (e.g. if there is an objective to manage for a specific habitat type) etc.

### Detectability of species

Even if an appropriate species is present at a site, we cannot expect to detect it perfectly within every plot. There is always a risk that one or more species is missed, regardless of whether we use survey plots or woodland walks. This risk should be acknowledged and accounted for. If our sampling design uses random representative, replicated plots it will allow us to do this by using well-established ecological methods ( I suggest [Occupancy Modelling techniques](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6362448/#:~:text=Multispecies%20occupancy%20models%20can%20estimate,while%20accounting%20for%20imperfect%20detection.) ideally, or [Species Accumulation Curves](https://en.wikipedia.org/wiki/Species_discovery_curve)).

Plot size, sampling effort etc will affect detection probability and the number of species expected to be recorded, even if they are present at the site. Available habitat area surrounding a plot will also influence the number of species expected to occupy it, but this effect is distinct from the impact of sampling-area on the number of species **detected** [@matthews2016].

# Respondent value functions and weights

```{r load curated data, echo = F, include = F, eval = T }
extraction.location <- "Data\\Delphi round 1\\"
load(paste0(extraction.location,"curated.RData"))
```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
# library(readxl)
library(tidyverse)
library(ggpubr)
library(plotly)
library(htmltools)

# library(stringr)
# library(plotly)
# library(htmlwidgets)
# library(gridExtra)
# library(grid)
# library(DT)
# library(ggiraph)
# library(cowplot)
# library(ggnewscale)
# library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

```

###  {.tabset .tabset-fade .tabset-pills}

```{r vf Canopy Nativness - init, echo = F , results="asis"}

ind.num = 4
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

```

`r paste(length(unique(filtered_data$respondant_name)))` respondents.

::: panel-tabset
#### Continuous Value function

```{r vf - continuous, echo = F , results="asis"}
continuous_vf_fig()

```

#### Categorised value function

```{r vf - categorical, echo = F , results="asis"}

categorised_vf_fig()

# ## table
# filtered_data %>%
#   group_by(measure) %>%
#   summarise(Mean = mean(value, na.rm = TRUE) %>% round(digits = 1),
#             'Standard deviation' = sd(value, na.rm = TRUE) %>% round(digits = 1)) %>% 
# knitr::kable()

```

#### Standardised weights

```{r std weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name, weight_variable = "weight_standardised")

```

#### Weights

```{r vf - weights, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)

```
:::

### Comments on data

VFs:

-   Feeling consensus may strengthen easily here

-   Particular focus useful for 1-2 species

-   Different ideas as to the shape of that relationship: linear[^08_native_tree_spceis_richness-1], vs sigmoid, S-shaped relationship[^08_native_tree_spceis_richness-2].

[^08_native_tree_spceis_richness-1]: e.g. Saul Herbert

[^08_native_tree_spceis_richness-2]: e.g. Liam Plummer

Weightings

-   Big variation, some making it least important, some making it most.

# Respondent comments

### Considerations for indicator design

Large-scale reasons for a geographic bias in species richness.

### Priority exceptions

Woodland type

-   Montane [^08_native_tree_spceis_richness-3] and upland woods: less expected[^08_native_tree_spceis_richness-4]

-   Wetwood[^08_native_tree_spceis_richness-5]

-   Planted woodland creation - may have a more even mix of species than expected elsewhere[^08_native_tree_spceis_richness-6], but natural colonisation may be dominated by pioneering species[^08_native_tree_spceis_richness-7]

-   Wood pasture[^08_native_tree_spceis_richness-8]

-   Coppice[^08_native_tree_spceis_richness-9]

[^08_native_tree_spceis_richness-3]: Kylie Jones Mattock

[^08_native_tree_spceis_richness-4]: Peter Lowe, Chris Reid, Kylie Jones Mattock, Liam Plummer

[^08_native_tree_spceis_richness-5]: Chris Reid, David Logan, Dean Kirkland, Karen Hornigold, Sonia Lorenzo-Martin

[^08_native_tree_spceis_richness-6]: Dave Bonsall

[^08_native_tree_spceis_richness-7]: Saul Herbert

[^08_native_tree_spceis_richness-8]: David Logan, Dean Kirkland, Nick Reed-Beale, Sonia Lorenzo-Martin

[^08_native_tree_spceis_richness-9]: Dave Bonsall, Peter Lowe, Bob Epsom

> Note: all these wood type exceptions could potentially be dealt with by using a suitable list of appropriate/expected species

Topography: flat, homogeneous ground will have lower expected tree species diversity[^08_native_tree_spceis_richness-10]

[^08_native_tree_spceis_richness-10]: Alasdair Firth

Different species may have different wildlife value e.g. oak[^08_native_tree_spceis_richness-11]

[^08_native_tree_spceis_richness-11]: Jim Smith-Wright

Functional traits: Presence/number of different functional groups of trees (e.g. nectar sources) may be more important than species number[^08_native_tree_spceis_richness-12]

[^08_native_tree_spceis_richness-12]: Jim Smith-Wright

Scale/landscape context: a small wood plot with a few species that are otherwise rare in the landscape, or provide other landscape-scale functioning[^08_native_tree_spceis_richness-13]

[^08_native_tree_spceis_richness-13]: Alasdair Firth

Avoid driving inappropriate species diversification e.g. :

-   Inappropriate supplementary planting within Ancient woodland sites to boost species richness[^08_native_tree_spceis_richness-14].

-   Where simplified species structure is part of management for target species/community[^08_native_tree_spceis_richness-15]

-   Where specific but relatively simplified habitat enhances biodiversity at larger scales[^08_native_tree_spceis_richness-16].

-   Where inappropriate for landscape/designation[^08_native_tree_spceis_richness-17]

[^08_native_tree_spceis_richness-14]: Andy Sharky

[^08_native_tree_spceis_richness-15]: Daniel Aitken

[^08_native_tree_spceis_richness-16]: Peter Lowe

[^08_native_tree_spceis_richness-17]: David Venables

### Lower priority exceptions

Arboretums[^08_native_tree_spceis_richness-18]

[^08_native_tree_spceis_richness-18]: Kylie Jones Mattock

### Concerns already covered by this or other indicators

Isolated woods where species pool is limited in landscape[^08_native_tree_spceis_richness-19]

[^08_native_tree_spceis_richness-19]: Bob Epsom

> Would just score poor, prompting considerations RE supplementary planting etc.

Plantations[^08_native_tree_spceis_richness-20]

[^08_native_tree_spceis_richness-20]: Kylie Jones Mattock

> Mono-cultures will just score poorly

# Other approaches

## Woodland Wildlife Toolkit

Record main tree and shrub species present in the upper canopy (\>5m) within plots

-   3: \>= 5 species

-   2: 3-4 species

-   1: 0-2 species

## NFI

1 ha plot count tree and shrub species combined

-   Favorable: \> 4 species

-   Intermediate: 3 - 4

-   Unfavorable: 0 - 2

## Current WT estate condition assessment

Number of trees and shrubs present (recorded separately)

# References
