# Canopy nattiness

## Justification

Invasive species negatively impact biodiversity.

## Measuring

DOMIN/percentage cover for plots for each invasive present.

Note maximum height of Cherry laurel or Rhododendron ponticum if present.

Note invasive tree regeneration (e.g. Western hemlock)[^09_invasives-1]

[^09_invasives-1]: Chris Reid

Note detection/non detection of other invasives during woodland walk, ideally mapping their location, and nearby records from e.g. NBN.

> Indicator will automatically score 0 if high priority invasives are present

# Respondent value functions and weights

```{r load curated data, echo = F, include = F, eval = T }
extraction.location <- "Data\\Delphi round 1\\"
load(paste0(extraction.location,"curated.RData"))
```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
# library(readxl)
library(tidyverse)
library(ggpubr)
library(plotly)
library(htmltools)

# library(stringr)
# library(plotly)
# library(htmlwidgets)
# library(gridExtra)
# library(grid)
# library(DT)
# library(ggiraph)
# library(cowplot)
# library(ggnewscale)
# library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

```

###  {.tabset .tabset-fade .tabset-pills}

```{r vf Canopy Nativness - init, echo = F , results="asis"}

ind.num = 5
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

```

`r paste(length(unique(filtered_data$respondant_name)))` respondents.

::: panel-tabset
#### Continuous Value function

```{r vf - continuous, echo = F , results="asis"}

continuous_vf_fig()

```

#### Standardised weights

```{r std weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name, weight_variable = "weight_standardised")

```

#### Weights

```{r vf - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)

```
:::

### Comments on data

VF:

-   A lot of variation. Worth considering Woodland Wildlife Toolkit approaches when revising estimates.

# Respondent comments

### Considerations for indicator design

Should drive early detection and eradication of invasives

-   There should be a rapid exponential drop in condition with presence of any invasive[^09_invasives-2]

-   A little may not have a big impact initially[^09_invasives-3], but much easier to deal with before it spreads[^09_invasives-4], growth (and so the potential for impact and cost of removal) is likely exponential[^09_invasives-5].

[^09_invasives-2]: Liam Plummer, Jim Smith-Wright

[^09_invasives-3]: Martin Hugi

[^09_invasives-4]: Ewan McHenry

[^09_invasives-5]: Saul Herbert

Location may be worth noting: beside watercourses or boundary etc may have management implications[^09_invasives-6]

[^09_invasives-6]: Hannah Patterson, David Logan

Note landscape-presence of invasives - relevant for management[^09_invasives-7]

[^09_invasives-7]: Hannah Patterson

### Priority exceptions

May consider species such as Gualtheria shallon equal consideration to Rhododendron, over laurel[^09_invasives-8]

[^09_invasives-8]: Jim Smith-Wright

### Lower priority exceptions

Species - Some may spread more and be more costly/difficult to remove[^09_invasives-9]

[^09_invasives-9]: Hannah Patterson

> Not really ecologically-relevant, just needs to be a

Aggressiveness of rhododendron and laurel depends on soil type[^09_invasives-10]

[^09_invasives-10]: Jim Smith-Wright

-   In SE cherry laurel may not spread[^09_invasives-11]

[^09_invasives-11]: Dave Bonsall

Woodland creation - Less important[^09_invasives-12]

[^09_invasives-12]: Liam Plummer

Non-plant invasives? Grey squirrel, mink etc.

### Concerns already covered by this or other indicators

Different invasives are worse[^09_invasives-13]

[^09_invasives-13]: Karen Hornigold

> indicator will score 0 if high priority species are present

Age/structure important[^09_invasives-14]

[^09_invasives-14]: Ewan McHenry, Jim Smith-Wright

> Indicator now records of height/flowering status

Appropriately managed species that are an acceptable part of the historic/designed landscape/park/garden[^09_invasives-15].

[^09_invasives-15]: Jim Smith-Wright, David Venables

> This may mean willingly accepting dropping ecological condition, as a trade-off to historic/amenity-value

Diverse, well functioning woodlands should be less invasive-susceptible.

> Should therefore have a higher score for this indicator

Where control/eradication is difficult/expensive[^09_invasives-16]

[^09_invasives-16]: Rhiannon Hoy, Karen Hornigold

> The indicator provides justification for the cost, if it is a big problem. Management cost will have little ecological-relevance (beyond impact on effectiveness of control).

# Other approaches

## Woodland Wildlife Toolkit & NFI

Percentage cover for species at whole woodland level

Presence of Rhododendron or Cherry laurel gives automatic poor condition

-   Favorable: no invasive species

-   Intermediate: Rhododendron and cherry laurel absent and other invasives \< 10 % cover

-   Unfavorable: Rhododendron or cherry laurel present or other invasives \> 10 % cover

Species considered:

::: two.columns
-   Hottentot-fig (*Carpobrotus edulis*)

-   Cotoneaster (*Cotoneaster spp.*)

-   Japanese knot weed (*Fallopia japonica*)

-   Shallon (*Gaultheria shallon*)

-   Giant rhubarb (*Gunnera manicata*)

-   Giant hogweed (*Heracleum mantegazzianum*)

-   Himalayan Balsam (*Impatiens glandulifera*)

-   Laurel (*Lauraceae*)

-   Rhododendron (*Rhododendron ponticum*)

-   Snowberry (*Symphoricarpos albus*)
:::

## Current WT estate condition assessment

# References
