# Veteran trees

## Justification

Veteran features provide diverse niches for wildlife species. Their continuity benefits wildlife species communities.

## Measuring

Extend plot to a 1 ha square (50m from center in each direction) and count AVTs (up to a maximum).

Where recording of AVTs is effective (i.e. an inventory has been done). This can be use instead of field survey - May encourage better recording on estate[^11_avts-1].

[^11_avts-1]: Saul Herbert

Note where AVTs make up the majority of all trees present. There may be too many for a long-term sustainability[^11_avts-2].

[^11_avts-2]: Alasdair Firth

# Respondent value functions and weights

```{r load curated data, echo = F, include = F, eval = T }
extraction.location <- "Data\\Delphi round 1\\"
load(paste0(extraction.location,"curated.RData"))
```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
# library(readxl)
library(tidyverse)
library(ggpubr)
library(plotly)
library(htmltools)

# library(stringr)
# library(plotly)
# library(htmlwidgets)
# library(gridExtra)
# library(grid)
# library(DT)
# library(ggiraph)
# library(cowplot)
# library(ggnewscale)
# library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

```

```{r vf Canopy Nativness - init, echo = F , results="asis"}

ind.num = 7
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

```

`r paste(length(unique(filtered_data$respondant_name)))` respondents.

::: panel-tabset
#### Continuous Value function

```{r vf - continuous, echo = F , results="asis"}

continuous_vf_fig()

```

#### Standardised weights

```{r std weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name, weight_variable = "weight_standardised")

```

#### Weights

```{r vf - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)

```
:::

### Comments on data

VFs:

-   Biggest disagreement is about where max should occur

Weights:

-   Massive variation, may need some evidence or guidance.

# Respondent comments

### Considerations for indicator design

Need to include coppice stools, but no agreed way to assess veteran status[^11_avts-3]

[^11_avts-3]: Dave Bonsall

Extending sample plot to 1 ha may be difficult, particularly in dense vegetation or steep ground.

Somewhat subjective definition of a veteran tree[^11_avts-4]

[^11_avts-4]: Jim Smith-Wright

Should record future potential veterans, particularly important where the site has current veterans[^11_avts-5]

[^11_avts-5]: Alasdair Firth, Rhiannon Hoy

> Research Q: Is there good evidence for AVT density needed for e.g. a diverse invert assemblage?

### Priority exceptions

Woodland type:

-   Wood pasture/parkland:

    -   Good condition would require more[^11_avts-6], perhaps more than eight[^11_avts-7]

    -   More important here[^11_avts-8]

-   Young woodland - Expect none[^11_avts-9], or just lower[^11_avts-10], and have lower importance weighting[^11_avts-11]

-   Upland/montane[^11_avts-12],

-   Coppice[^11_avts-13]

-   Alder carr, wet willow woodland more dynamic woods - retention of trees to veteran is less important[^11_avts-14]

-   Upland birch: may be solely veteran[^11_avts-15], but also potentially very dynamic[^11_avts-16]

[^11_avts-6]: Chris Reid, Liam Plummer, Adam Thorogood, Rhiannon Hoy

[^11_avts-7]: Chris Reid

[^11_avts-8]: Liam Plummer, Jim Smith-Wright, Karen Hornigold

[^11_avts-9]: Chris Reid, Dave Bonsall, Kylie Jones Mattock, Hannah Patterson, Ian Moss, Daniel Aitken, David Logan, David Venables

[^11_avts-10]: Rhiannon Hoy, Sonia Lorenzo-Martin

[^11_avts-11]: Liam Plummer, Hannah Patterson, Louise Hackett, Karen Hornigold

[^11_avts-12]: Peter Lowe

[^11_avts-13]: Peter Lowe, Dave Bonsall

[^11_avts-14]: Adam Thorogood

[^11_avts-15]: Peter Lowe

[^11_avts-16]: Adam Thorogood

### Lower priority exceptions

Some species more value than others[^11_avts-17]

[^11_avts-17]: Jim Smith-Wright

Lowland ancient woodland often have no AVTs[^11_avts-18]

[^11_avts-18]: Peter Lowe

> Should they be managed with AVTs more in mind?

### Concerns already covered by this or other indicators

Commercial plantations[^11_avts-19]

[^11_avts-19]: Kylie Jones Mattock

> Will score low for this indicator if no AVTs present

If whole wood is old there is no space for young trees[^11_avts-20]

[^11_avts-20]: Alasdair Firth

> Should be dealt with via tree age or distribution and regeneration. Have Added suggestion to note where AVTs are the majority of trees present[^11_avts-21]

[^11_avts-21]: Ewan McHenry

High conservation value sites may have no veterans[^11_avts-22]

[^11_avts-22]: Daniel Aitken

> They would be scoring well for other indicators

# Other approaches

## Woodland Wildlife Toolkit

Map locations of veteran trees encountered during woodland walk and any other known.

Veteran defined based on species-specific DBH and/or \>=3 veteran features:

> Veteran features: major trunk cavities or hollowing; water pools in crevices; decay-caused small holes in trunk/larger branches/large roots; missing or loose bark; large quantities of canopy deadwood; sap seeping through bark; crevices sheltered from rainfall; fungi on trunk and larger branches; plants (not mosses or lichen) growing on trunk or branches.

-   Value 3: \> 4 per ha

-   Value 2: 1-2 per ha

-   Value 1: 0 AVTs

## NFI

Recorded within 1 ha plot:

-   Favorable: \> 2 per ha

-   Intermediate: 0.05 - 2 per

-   Unfavorable: \< 0.05 per ha

## Current WT estate condition assessment

# References
