# Habitat extent

## Justification

A plot (of a fixed size) within a larger woodland habitat area will function better ecologically. More surrounding habitat provides a greater pool of potential colonisers and more room for dynamism (also see Habitat Area Hypothesis[^12_extent-1]).

[^12_extent-1]: The habitat amount hypothesis states that the total amount of habitat in a landscape determines its species richness [@fahrig2013]

## Measuring

Area of woodland habitat patch size from GIS

> Alternative 1: percentage cover of woodland habitat within some distance from plots (500 m and/or 1 km?) may me more ecologically relevant.
>
> Alternative 2: Equivalent Connected Area of habitat (ecological functional connectivity metric).
>
> Q: Should other semi-natural habitat be considered?

# Respondent value functions and weights

```{r load curated data, echo = F, include = F, eval = T }
extraction.location <- "Data\\Delphi round 1\\"
load(paste0(extraction.location,"curated.RData"))
```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
# library(readxl)
library(tidyverse)
library(ggpubr)
library(plotly)
library(htmltools)

# library(stringr)
# library(plotly)
# library(htmlwidgets)
# library(gridExtra)
# library(grid)
# library(DT)
# library(ggiraph)
# library(cowplot)
# library(ggnewscale)
# library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

```

```{r vf Canopy Nativness - init, echo = F , results="asis"}

ind.num = 8
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]

filtered_data <- df[df$indicator_name == ind.matcher.df$indicator_name_for_extraction[ind.num], ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

```

`r paste(length(unique(filtered_data$respondant_name)))` respondents.

::: panel-tabset
#### Continuous Value function

```{r vf - continuous, echo = F , results="asis"}

continuous_vf_fig()

```

#### Weights

```{r vf - weights, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)

```

#### Standardised weights

```{r std weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name, weight_variable = "weight_standardised")

```

Weights, standardised so that all respondents varied the same amount
:::

# Respondent comments

### Considerations for indicator design

It is unclear what constitutes a "site" or "patch". Will a 20m gap mean a new "patch"? Does a different woodland type mean a different patch?

> Alternative 1 (percentage cover of woodland habitat within some distance) or 2 (ecological functional connectivity metric) would deal with this.

Should not undermine contribution of non-wooded priority habitat[^12_extent-2], e.g. by driving planting on it[^12_extent-3]

[^12_extent-2]: Hannah Patterson

[^12_extent-3]: Kylie Jones Mattock

Feeling that 50 ha is too small a maximum[^12_extent-4]

[^12_extent-4]: Peter Lowe, Alasdair Firth

This may be best viewed as "Landscape context"[^12_extent-5]

[^12_extent-5]: Richard Brown

### Priority exceptions

Shape, connectivity, patch size, edge area and landscape structure will interact with this[^12_extent-6]

[^12_extent-6]: Chris Reid, Kylie Jones Mattock, Liam Plummer, Alasdair Firth, Bob Epsom, David Venables

> Alternative 2 (ecological functional connectivity metric) would deal with this.

Landscape:

-   Where a rare woodland type gives an important contribution to the landscape, especially as part of a mosaic[^12_extent-7]
    -   E.g. a small dense grove with limited structure and no veteran trees may be less important within a landscape that is rich in wood pastures with lots of veteran trees. Equally, expanding that patch into the woodpasture may be inappropriate[^12_extent-8].

    -   <div>

        > Including that woodpasture within the "extent" might solve this?

        </div>
-   Importance greater in landscapes with low habitat: more is "required" of the site to sustain itself[^12_extent-9].
-   Wet woodland - often small and OK condition[^12_extent-10].

[^12_extent-7]: Saul Herbert, David Logan

[^12_extent-8]: Ewan McHenry

[^12_extent-9]: Liam Plummer

[^12_extent-10]: Dean Kirkland, David Logan

Small woods within high-quality open habitat mosaic or upland peat[^12_extent-11]

[^12_extent-11]: Kylie Jones Mattock, David Logan

> Could this be dealt with by measuring percentage woodland cover within some relatively limited dispersal distance e.g. 500m?

Higher importance where woodland has been lost in the landscape in recent years[^12_extent-12]

[^12_extent-12]: Rhiannon Hoy

### Lower priority exceptions

Small woods in farmland woods, hedges and edges[^12_extent-13] or urban woods surrounded and unable to expand. Might be stuck in "Poor" condition because of this indicator: suggested lower importance here[^12_extent-14]

[^12_extent-13]: Kylie Jones Mattock

[^12_extent-14]: Hannah Patterson

> Not ecologically relevant. These would just score poorly for this particular indicator and need to be able to attain good condition based on other indicators.

Large monocultures e.g. Sitka plantations[^12_extent-15]

[^12_extent-15]: Hannah Patterson

> May score well for this indicator (unless "semi-natural habitat" extent is specified), but should score terribly on other indicators. Important to balance weightings accordingly.

Small woods can be important temporal refugia for species that are out competed in more connected woodland[^12_extent-16]

[^12_extent-16]: Ewan McHenry, Jim Smith-Wright

> Q: When does a wood becomes so small that it wouldn't be reasonable to see all indicators well represented? - Or when it could be a component of indicators in a landscape?[^12_extent-17]

[^12_extent-17]: Saul Herbert

## Woodland Wildlife Toolkit & NFI

Size of NFI GB Woodland Base Map parcel

Value 3: \>20 ha

Value 2: 5 - 10 ha

Value 1: \< 5 ha

## Current WT estate condition assessment

# References
