# Habitat extent

## Justification

A plot (of a fixed size) within a larger woodland habitat area will function better ecologically. More surrounding habitat provides a greater pool of potential colonisers and more room for dynamism (also see Habitat Area Hypothesis[^12_extent-1]).

[^12_extent-1]: The habitat amount hypothesis states that the total amount of habitat in a landscape determines its species richness [@fahrig2013]

## Measuring

Area of woodland habitat patch size from GIS

> Alternative 1: percentage cover of woodland habitat within some distance from plots (500 m and/or 1 km?) may me more ecologically relevant.
>
> Alternative 2: Equivalent Connected Area of habitat (ecological functional connectivity metric).
>
> Q: Should other semi-natural habitat be considered?

# Respondent value functions and weights

```{r load curated data, echo = F, include = F, eval = T }
extraction.location <- "Data\\Delphi round 1\\"
load(paste0(extraction.location,"curated.RData"))
```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
# library(readxl)
library(tidyverse)
library(ggpubr)
library(plotly)
library(htmltools)

# library(stringr)
# library(plotly)
# library(htmlwidgets)
# library(gridExtra)
# library(grid)
# library(DT)
# library(ggiraph)
# library(cowplot)
# library(ggnewscale)
# library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

```

```{r vf Canopy Nativness - init, echo = F , results="asis"}

ind.num = 8
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]

filtered_data <- df[df$indicator_name == ind.matcher.df$indicator_name_for_extraction[ind.num], ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

```

`r paste(length(unique(filtered_data$respondant_name)))` respondents.

::: panel-tabset
#### Continuous Value function

```{r vf - continuous, echo = F , results="asis"}
#| column: page

continuous_vf_fig()

```

#### Weights

```{r vf - weights, echo = F , results="asis"}
#| column: page

ggplot_resp_weight_ind(indicator_name = indicator_name)

```
:::

### Comments on data

VFs:\
- consensus may form

Weights:\
- Some think it most important, some more middling

# Respondent comments

### Considerations for indicator design

It is unclear what constitutes a "site" or "patch". Will a 20m gap mean a new "patch"? Does a different woodland type mean a different patch?

> percentage habitat

Should not undermine contribution of non-wooded priority habitat[^12_extent-2], e.g. by driving planting on it[^12_extent-3]

[^12_extent-2]: \[\[Hannah Patterson\]\]

[^12_extent-3]: \[\[Kylie Jones Mattock\]\]

Feeling that 50 ha is too small a maximum[^12_extent-4]

[^12_extent-4]: \[\[Peter Lowe\]\], \[\[Alasdair Firth\]\]

This may be best viewed as "Landscape context"[^12_extent-5]

[^12_extent-5]: \[\[Richard Brown\]\]

### Priority exceptions

Shape, connectivity and landscape structure will interact with this[^12_extent-6]

[^12_extent-6]: \[\[Chris Reid\]\], \[\[Kylie Jones Mattock\]\], \[\[Liam Plummer\]\], \[\[Alasdair Firth\]\]

landscape:

-   Where a rare woodland type gives an important contribution to the landscape, especially as part of a mosaic[^12_extent-7]
    -   E.g. a small dense grove with limited structure and no veteran trees may be less important within a landscape that is rich in wood pastures with lots of veteran trees. Equally, expanding that patch into the woodpasture may be inappropriate[^12_extent-8].

    -   <div>

        > Perhaps including that woodpasture within the "extent" would solve this?

        </div>
-   Importance greater in landscapes with low habitat: more is "required" of the site to sustain itself[^12_extent-9].

[^12_extent-7]: \[\[Saul Herbert\]\]

[^12_extent-8]: Ewan McHenry

[^12_extent-9]: \[\[Liam Plummer\]\]

Urban woods surrounded and unable to expand, these might be constantly stuck in "Poor" condition because of this indicator: suggested lower importance here[^12_extent-10]

[^12_extent-10]: \[\[Hannah Patterson\]\]

> They need to be able to attain good condition based on other indicators.

Small woods within high-quality open habitat mosaic or upland peat[^12_extent-11]

[^12_extent-11]: \[\[Kylie Jones Mattock\]\]

> Could this be dealt with by measuring percentage woodland cover within some relatively limited dispersal distance e.g. 500m?

### Lower priority exceptions

Farmland woods, hedges and edges[^12_extent-12]

[^12_extent-12]: \[\[Kylie Jones Mattock\]\]

> These would just score poorly for this particular indicator[^12_extent-13]

[^12_extent-13]: \[\[Ewan McHenry\]\]

Large monocultures e.g. Sitka plantations[^12_extent-14]

[^12_extent-14]: \[\[Hannah Patterson\]\]

> May score well for this indicator (unless "semi-natural habitat" extent is specified), but should score terribly on other indicators. Important to balance weightings accordingly.

Small woods can be important temporal refugia for species that are out competed in more connected woodland[^12_extent-15]

[^12_extent-15]: \[\[Ewan McHenry\]\], \[\[Jim Smith-Wright\]\]

> Q: When does a wood becomes so small that it wouldn't be reasonable to see all indicators well represented? - Or when it could be a component of indicators in a landscape?[^12_extent-16]

[^12_extent-16]: \[\[Saul Herbert\]\]

# Other approaches

## Woodland Wildlife Toolkit & NFI

Size of NFI GB Woodland Base Map parcel

Value 3: \>20 ha

Value 2: 5 - 10 ha

Value 1: \< 5 ha

## Current WT estate condition assessment

# References
