# Human damage

## Justification

High pollution reduces species diversity and natural processes[^18_anthro-damage-1]

[^18_anthro-damage-1]: Alasdair Firth

## Measuring

This indicator requires significant development

Attempting to capture:

::: two-columns
-   Anthropogenic damage

-   Soil damage

-   Compaction, erosion

-   Farming spill over: fertiliser, spraying & pesticide

-   Livestock feeding in wood with potential to cause damage

-   Dumping

-   N+ indicator plants
:::

May be better to separate out damage based on spatial scale, intensity and permanence.

# Respondent value functions and weights

```{r load curated data, echo = F, include = F, eval = T }
extraction.location <- "Data\\Delphi round 1\\"
load(paste0(extraction.location,"curated.RData"))
```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
# library(readxl)
library(tidyverse)
library(ggpubr)
library(plotly)
library(htmltools)

# library(stringr)
# library(plotly)
# library(htmlwidgets)
# library(gridExtra)
# library(grid)
# library(DT)
# library(ggiraph)
# library(cowplot)
# library(ggnewscale)
# library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

```

```{r vf Canopy Nativness - init, echo = F , results="asis"}

ind.num = 14
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]

filtered_data <- df[df$indicator_name == ind.matcher.df$indicator_name_for_extraction[ind.num], ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

```

`r paste(length(unique(filtered_data$respondant_name)))` respondents.

::: panel-tabset
#### Continuous Value function

```{r vf - continuous, echo = F , results="asis"}

continuous_vf_fig()

```

#### Standardised weights

```{r std weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name, weight_variable = "weight_standardised")

```

#### Weights

```{r vf - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)

```
:::

### Comments on data

VFs: Big range, but may converge to a consensus.

-   Recommend referring to the [Woodland Wildlife Toolkit] methods to contextualize

Weights:

-   Full range

# Respondent comments

### Considerations for indicator design

Fertilizer spraying and wash-off from adjacent land not always obvious[^18_anthro-damage-2]

[^18_anthro-damage-2]: Richard Brown

May require soil/water samples

Should the source of damage be identified and recorded?

-   Impacts may be subtle and take years to develop
-   Onsite vs offsite sources

OPAL has produced info to help nitrogen-loving lichen ID as part of their [Air Quality Monitoring](https://www.imperial.ac.uk/media/imperial-college/research-centres-and-groups/opal/AIR-16pp-booklet_legacy.pdf)

Could record desire lines[^18_anthro-damage-3].

[^18_anthro-damage-3]: Rhiannon Hoy

### Priority exceptions

Depends a lot on type of damage[^18_anthro-damage-4], some human impacts not necessarily negative[^18_anthro-damage-5]

[^18_anthro-damage-4]: Jim Smith-Wright, Hannah Patterson, Ian Moss, Martin Hugi, Louise Hackett, David Venables

[^18_anthro-damage-5]: Kylie Jones Mattock

Nutrient enrichment indicator plants may indicate land use before planting, not ongoing threat. E.g. Planting on arable field[^18_anthro-damage-6]

[^18_anthro-damage-6]: Richard Brown

Ongoing/one-off/accumulating instances[^18_anthro-damage-7]

[^18_anthro-damage-7]: Hannah Patterson, Louise Hackett

"Damage" caused by woodland management operations that benefit long term conservation objectives[^18_anthro-damage-8].

[^18_anthro-damage-8]: Louise Hackett

### Lower priority exceptions

Commercial objectives[^18_anthro-damage-9]

[^18_anthro-damage-9]: Kylie Jones Mattock

Secondary woodlands[^18_anthro-damage-10]

[^18_anthro-damage-10]: Kylie Jones Mattock

Soil disturbance may be good for natural regeneration[^18_anthro-damage-11] and ecological functioning

[^18_anthro-damage-11]: Ian Moss

Urban/amenity/higher access woods - may be relatively good, compared to similar sites, but have more damage than rural[^18_anthro-damage-12]

[^18_anthro-damage-12]: David Logan, David Venables

> Don't think this is ecologically relevant. We might expect that urban woods have more damage and score poorly, should make like-for-like comparisons, but it is useful to be able to say e.g.: "urban woods have X% more damage.

### Concerns already covered by this or other indicators

Edge may be affected more[^18_anthro-damage-13]

[^18_anthro-damage-13]: Alasdair Firth

> May just score worse. Could provide evidence for e.g. campaigning around buffering.

Demographics of neighboring community, perception of woodland culture[^18_anthro-damage-14]

[^18_anthro-damage-14]: David Venables

> Not ecologically relevant, insofar as the same damage being valued differently.

# Other approaches

## Woodland Wildlife Toolkit

On woodland walk, record evidence of nutrient enrichment and/or damaged ground, including patches of nettle and/or goose grass/cleavers (nutrient enrichment indicator) \>0.1 ha

-   Favorable: No nutrient enrichment or damaged ground

-   Intermediate: \< 1 ha total nutrient enrichment and/or \< 20% damaged soil

-   Poor: \> 1 ha total nutrient enrichment and/or \> 20% damaged soil

> Note: the bigger the woodland, the more likely that damage/enrichment will be found and be \> 1 ha, purely because a larger area is searched. This would be better expressed as a % of the wood.
>
> Note: "and/or" wording in poor class is ambiguous. should be "or"

Note soil damage: - Deep ruts - Excessive/continuous compaction

## Current WT estate condition assessment

Human impacts, split by one off/continuous

# References
