---
title: "Assessing Woodland Ecological Condition"
author: "Ewan McHenry"
date:    |
  `r format(Sys.time(), '%d %B, %Y')`
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cosmo
---
<style>
.html-widget {
  max-width: 1000px;
  margin: auto;
}
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=5, fig.path='Figs/',
                      table.width = "50%",
                      echo=FALSE, warning=FALSE, message=FALSE)
forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

```

# Introduction  

```{r data_extraction_from_excelsheets - preliminary work, echo = F, include = F, eval = F }

# extract from excel sheets ----
source("Scripts\\Extract_expert_info.R")
# if all is well this should:
# save ind.matcher ind.matcher.df and expert.data (expert opinion responses from the Delphi forms) 
# as files within the specified extraction.location to be loaded together

# curation ----
source("Scripts\\curation of extracted data.R")
# this one loads the data and does some curation, scaling responses etc
# then saves the dataframes

# then later I loads it (in Italian-American mafiosi accent)

```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
library(readxl)
library(tidyverse)
library(stringr)
library(plotly)
library(ggpubr)
library(mgcv)
library(RColorBrewer)
library(htmltools)
library(htmlwidgets)
library(gridExtra)
library(grid)
library(DT)
library(ggiraph)
library(cowplot)
library(ggnewscale)
library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

load(paste0(extraction.location,"curated.RData"))
```


```{r respondant colours, echo = F, include = F, eval = T }
respondant_colours <- glasbey.colors(nlevels(respondants_with_something)+1)[-1] %>% as.character() %>% 
    setNames(., respondants_with_something)

```


This document explores the expert opinion gathered from staff at the Woodland Trust, related to the measurement of `r has_completed$indicator_name %>% unique() %>% length()` indicators of woodland ecological condition, and the relationship between these indicators' and a wood's potential to promote flourishing wildlife species communities (by estimating "value functions" and importance weightings for each indicator). 

Expert opinion on the value function and weight for each indicator was elicited using the Delphi method. The Delphi method is designed to collect and distill expert knowledge, using repeated surveys to refine estimates, increasing their reliability 
<!-- (Edwards et al., 2011) -->
and identifying where agreements and disagreements in understanding occur.   
<!-- (Buckley, 1995; Kangas, Alho, Kolehmainen, & Mononen, 1998; Landeta, 2006). --> 
  
The method relies on repetitive surveying in at least two to three ‘rounds’ where participants are asked the same questions. After each round participants review and consider the answers given by all panel-members, and are then given the opportunity to revise their answers. The Delphi method has a history of successful use in supporting conservation action where empirical evidence is insufficient and information is needed rapidly to inform decisions. It's transparency and repeatability also add to its appeal.  

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName">Indicators of Woodland Ecological Condition - Show/Hide</button>  
<div id="BlockName" class="collapse">  

```{r  indicators table, echo = F }

ind.table <- ind.matcher.df [,1:2]
names(ind.table) <- c("Indicator", "Description")
ind.table %>% 
   datatable(#extensions = c('FixedColumns',"FixedHeader"),
              options = list(scrollX = T, 
                             paging = F,
                             fixedHeader=F,
                             dom = 't',
                             fixedColumns = list(leftColumns = 1) #,
                             # buttons = c('copy', 'csv')
                             ),
              rownames = F#,
              #container = sketch.lcm.areas
              ) %>% 
    formatStyle( columns = c("Indicator"), fontWeight = 'bold', `text-align` = 'left', "white-space" = "nowrap") 

```
</div>

# Methods  

## Ellicitation of expert opinion

Each participant either attended an introductory workshop to the process, or was taken through individually. A sub-section of participants were also involved in development the proposed methods to measure each indicator in the field. Answers were submitted into individual Excel spread sheets, with questions relating to:  

- Value functions describing how each indicator related to WEC, in typical conditions.  
  - These were entered as points representing potential indicator measurements and that potential measurement's corresponding relative value (on a 0-100 scale from low to high).
- Weightings describing the relative importance of each indicator to WEC, in typical conditions (by which an indicator's value would be multiplied).
- Certainty around value function and weighting estimates (subjective score 0-5), to aid participants in reviewing responses.
- Comments or concerns relating to: situations where these "typical" responses might be inappropriate, the proposed method to measure indicators, and any other relevant information.  

Participants were encouraged to keep a clear distinction between the quantitative responses provided (i.e. value functions and weightings under typical conditions) and exceptions to those typical relationships (that they highlighted in their comments).  

Participants were instructed that noted exceptions would be analysed and, where appropriate, used to inform future development of the WEC measure to account for important nuance.

## Analysis Method
### Scaling values and weights

Values and weights are all measured relative to each other. Some respondents did not use the full 0 to 100 range for all value functions and weights. Comparison between participants requires that they all use the same scale, and so all value functions where scaled from 0 to 100 and weightings were scaled so that the highest indicator weighting for each panelist was 100.

<button class="btn btn-primary" data-toggle="collapse" data-target="#Scaling_info">Details of responses requiring scaling - Show/Hide</button>  
<div id="Scaling_info" class="collapse">  


**Value functions** scaled for:  
`r paste0("<br>- ", which.scaled$vf)`  

**Weights** were scaled for:  

`r paste0("<br>- ", which.scaled$wt)`  
</div>


### Edits to responses

In some cases where deemed appropriate (e.g. where respondents had described a linear trend using few, widely spaced points), points were added to value functions by visual interpolation. Some points where also added by extrapolation where a maximum/minimum value was clearly implied, and corrections where made to clearly erroneous points following individual consultation with respondents. 



## Reviewing results between rounds

A copy of this report was circulated to respondents between rounds to allow respondents to investigate estimates provided by others, and their certainties.

Where appropriate value functions where plotted categorically (with box plots summarising the estimated value at each category), and continuously. Indicator relative weightings were presented on boxplots, including the mean and median estimates. Interactive hover-text displayed respondent name and their certainty in their estimate.


# Survey completion

## Round 1 {.tabset .tabset-fade .tabset-pills}
`r has_completed$respondent_names %>% unique() %>% length()` expert practitioners from Woodland Trust and Plantlife staff were invited to provide their opinion the value and weight of indicators.  
`r sum(completed_summary[,2:5] %>% rowSums() >0)` have provided a response as of `r date()`.   
Of these, `r sum((completed_summary[,2:5] %>% rowSums() >0) & (completed_summary[,2:5] %>% rowSums() < (4* has_completed$indicator_name %>% unique() %>% length() ) ))` remain partially incomplete.  

<button class="btn btn-primary" data-toggle="collapse" data-target="#completion_summary">Survey completion summary - Show/Hide</button>  
<div id="completion_summary" class="collapse">  

The chart below summarises participant completion of different elements of the survey. For more detail see table (button below chart).

``` {r  participation_chart - make, echo = F, eval = F }

# use df to ggplot
# The first legitimate use of a pie chart in my carer!

max_count = max(completed_summary$vf)

create_gg_pie_chart <- function(count, max_count) {
  percentage <- count / max_count * 100
  pie_chart <- ggplot(data.frame(x = c("Done", "Remaining"), y = c(count, max_count - count)),
                      aes(x = "", y = y, fill = x)) +
    geom_bar(width = 1, stat = "identity") +
    scale_fill_manual(values = c("steelblue", "lightgray")) +  # Specify custom colors
    coord_polar(theta = "y") +
    theme_void() +
    theme(legend.position = "none")
  
  return(pie_chart)
  
  pie_matrix <- arrangeGrob(grobs = gg_pie_charts, nrow = length(completed_summary$respondent_names), ncol = length(colnames(completed_summary[, c("wt", "wt_cert", "vf", "vf_cert")])))
# Add row names (respondent names)
row_names <- completed_summary$respondent_names

# Add column names
col_names <- c("Value func", "VF cert", "Weights", "wt cert")

combine <- rbind(tableGrob(t(c(col_names)), theme = ttheme_minimal(), rows = ""), 
                 cbind(tableGrob(row_names, theme = ttheme_minimal()), 
                       arrangeGrob(grobs = gg_pie_charts, nrow = length(row_names), ncol = length(col_names)),  size = "last"), size = "last")
}

# gg_pie_charts <- list(NA)
# bk = 1
# for (i in 1:(length(completed_summary$respondent_names))){
#   for (ii in 1:(dim(completed_summary)[2]-1)){
#     count <- completed_summary[i,1+ii] %>% as.numeric()
#     gg_pie_charts[[bk]] <-  create_gg_pie_chart(count, length(unique(df$indicator_name)))
#   bk = bk+1
#     }}
#   
save(combine, file = "Figs//gg_pie_charts.RData")

```


``` {r  participation_chart - plot, echo = F }
load(file = "Figs//gg_pie_charts.RData")

# Convert the list of pie charts to a grid using grid.arrange

grid.newpage()
grid.draw(combine)

```
Figure 1. Completion of expert opinion survey components by each panelist (coloured portion of pies) for value functions ("Value func"), value function certainties ("VF cert"), indicator weightings ("Weights") and weighting certainties ("wt certs").  

</div>
  
<button class="btn btn-primary" data-toggle="collapse" data-target="#completion_details">Survey completion details - Show/Hide</button>  
<div id="completion_details" class="collapse">  

```{r what needs done by participants details, eval = T}

# Create a data frame with participant names and the four columns
participant_table <- data.frame(respondent_names = completed_summary$respondent_names,
                                vf = "",
                                vf_cert = "",wt = "",
                                wt_cert = "",
 stringsAsFactors = FALSE)

# Iterate over each participant and update the table with the indicators not completed
for (i in 1:nrow(completed_summary)) {
  n.indicators <- has_completed$indicator_name %>% unique() %>% length()
  participant <- completed_summary[i, "respondent_names"] %>% as.character()
  not_completed_wt <- has_completed$sheet_name[has_completed$respondent_names == participant & has_completed$wt == FALSE]
  not_completed_wt_cert <- has_completed$sheet_name[has_completed$respondent_names == participant & has_completed$wt_cert == FALSE]
  not_completed_vf <- has_completed$sheet_name[has_completed$respondent_names == participant & has_completed$vf == FALSE]
  not_completed_vf_cert <- has_completed$sheet_name[has_completed$respondent_names == participant & has_completed$vf_cert == FALSE]
  
  if (length(not_completed_wt) == n.indicators) not_completed_wt <- "all"
  if (length(not_completed_wt_cert) == n.indicators) not_completed_wt_cert <- "all"
  if (length(not_completed_vf) == n.indicators) not_completed_vf <- "all"
  if (length(not_completed_vf_cert) == n.indicators) not_completed_vf_cert <- "all"

  participant_table[participant_table$respondent_names == participant, "wt"] <- paste(not_completed_wt, collapse = ", ")
  participant_table[participant_table$respondent_names == participant, "wt_cert"] <- paste(not_completed_wt_cert, collapse = ", ")
  participant_table[participant_table$respondent_names == participant, "vf"] <- paste(not_completed_vf, collapse = ", ")
  participant_table[participant_table$respondent_names == participant, "vf_cert"] <- paste(not_completed_vf_cert, collapse = ", ")
}

# Print the participant table

names(participant_table) <- c("Respondent", "Value Functions", "vf certs", "Weights", "wt certs")

participant_table %>%  
    datatable(extensions = c('FixedColumns',"FixedHeader"),
              options = list(scrollX = T, 
                             scrollY = '400px',  # Set the desired height for the table
                             scrollCollapse = TRUE,

                             paging = F,
                             fixedHeader=F,
                             dom = 't',
                             autoWidth = TRUE, fixedColumns = list(leftColumns = 1) #,
                             # fixedColumns = list(leftColumns = 1) ,
                             # buttons = c('copy', 'csv')
                             ),
              rownames = F#,
              #container = sketch.lcm.areas
              ) %>% 
    formatStyle( columns = c("Respondent"), fontWeight = 'bold', `text-align` = 'left', "white-space" = "nowrap")# %>% 

```
</div>

# Indicatior weights

## Raw weights of all indicators

```{r raw weights, echo = F , results="asis"}
#-----------------------------------------------------
# PLOT WEIGHTS --------------------------------------------------
#-----------------------------------------------------

plot <- ggplot() +
  geom_boxplot(data = just.one.df, aes(x = sheet_name, y = weight), 
               width=0.3, color="grey70", fill = "white", alpha=0.2) +
  geom_point(data = just.one.df, 
             position = position_jitter(width = 0.2, height = 0),
             size = 1, shape = 16,
             aes(x = sheet_name, y = weight, color = respondant_name,
                 text = map(
                   paste0("<b>", `respondant_name`, "</b><br>",
                          "<b>Weight: </b>", weight, " (certainty ", cert_weight,")<br>"),
                   HTML))) +
  stat_summary(data = just.one.df, aes(x = sheet_name, y = weight,
                                       text = map(
                                         paste0("Mean weight"),
                                         HTML)),
               geom = "point", fun = mean,
               color = "red", shape = 3, size = 2,
               position = position_nudge(x = 0.05, y = 0)
               ) +
  labs(x = NULL, y = "Relative importance (weight)", 
       title = "Indicator weights",
       color = "Respondant") +
  theme_pubr()+
  theme(plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 12),
        legend.title = element_text(size = 12 , face = "bold"),
        legend.text =  element_text(size = 12),
        axis.text.x = element_text(angle = -60, hjust = 0, size = 10),
        axis.text.y = element_text(size = 12),
        axis.title =  element_text(size = 12),
        legend.position = "right",
        plot.margin = margin(0.5, 1, 0.5, 1, "cm")
  ) 
ggplotly(plot, tooltip = "text", dynamicTicks = F) %>% 
  config(displayModeBar = F) %>% 
  layout(yaxis = list(range = c(-5, 105)))

```

## Correlation between weights

Correlation plot of indicators below.  
- High positive correlation (large & blue) means that these indicators similarly weighted by individuals respondents (i.e. both high/medium/low).  
- High negative correlation (large & red) indicators are consistently given different weights by individuals respondents (i.e. one high and one low).
- Low correlation (small) indicators have little relationship  weights by individuals respondents (i.e. one high and one low).


```{r corr weights, echo = F , results="asis"}

# Correspondants analysis of respondants by weights given to different indicators
library("FactoMineR")
library("factoextra")

for.ca <- just.one.df %>%
  dplyr::select(sheet_name, 
                respondant_name, weight) %>%
  spread(key = sheet_name, value = weight) 
#make respondant name the row names
rownames(for.ca) <- for.ca$respondant_name
for.ca <- for.ca[, -1]
colnames(for.ca) <- c ("Age", "Native", "Vert.", "N spp.", "Invas.", "Deadwood",
                       "Vet.", "Ext.", "Regen", "Herb.", "Health", "Flora", "Horiz.",
                       "Anthro.")

# make corrplot of the different indicators
library("corrplot")

for.ca %>% na.omit() %>% 
  cor() %>% 
  corrplot(method = "circle", diag = F, type = "upper")


```



# Reponses by indicator  

## Tree age/size distribution 
### {.tabset .tabset-fade .tabset-pills}
```{r vf Tree age size distribution - init, echo = F , results="asis"}

ind.num = 1
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Categorised value function
```{r vf Tree age size distribution - categorical, echo = F , results="asis"}


categorised_vf_fig()

## table
filtered_data %>%
  group_by(measure) %>%
  summarise(Mean = mean(value, na.rm = TRUE) %>% round(digits = 1),
            'Standard deviation' = sd(value, na.rm = TRUE) %>% round(digits = 1)) %>% 
knitr::kable()

```

#### Continuous Value function
```{r vf Tree age size distribution - continuous, echo = F , results="asis"}


continuous_vf_fig <- function(line.col = NA){
  plot <- ggplot_gam_resp_vf(indicator_name = indicator_name,
                           x.lab = ind.axis.title, gam.col = line.col)

ggsave(filename = paste0("Figs//ind_",
                         formatC(ind.num, width = 2, format = "d", flag = "0"),
                         "_", indicator_name, "vf_continuous.png"),
       plot = plot,
       width = 300, height = 150, units = "mm")

ggplotly(plot, tooltip = "text", dynamicTicks = F) %>% 
    config(displayModeBar = F) %>% 
    layout(yaxis = list(range = c(-5, 105)))
}
  
continuous_vf_fig()

```

#### Weights
```{r vf Tree age size distribution - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)

```

### Comments

VFs - consensus  likely:  
- All continuous relationships have same shape.  
- Complete agreement on upper and lower value.  
- Some put maximum value at 3 classes.  

Weight - Disagreement:  
- Two camps, giving relatively low and high weights (compared to other indicators) at c. 60 and 90, respectively.  


## Canopy nativness 
### {.tabset .tabset-fade .tabset-pills}

```{r vf Canopy Nativness - init, echo = F , results="asis"}

ind.num = 2
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Canopy Nativness - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Canopy Nativness - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments  

VF:  
- Similar general trends, but a lot of variation
- Different shapes - Some linear, some S-shaped, some exponential, some logarithmic("r" shaped)
- disagreement RE max value  

Recommend focus for thought on:  
- lower 25% native 
- upper 75-100%


Weightings: 
- fairly wide spread
- 5 or so think it amongst most important
- rest think it is middle-importance
 

## Vertical Structure 

Of four possible categories: >15m; 5-15m; 2-5m; 0.5-2m; measured at plot level

### {.tabset .tabset-fade .tabset-pills}

```{r vf Vertical Structure - init, echo = F , results="asis"}

ind.num = 3
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Categorised value function
```{r vf Vertical Structure - categorical, echo = F , results="asis"}
categorised_vf_fig()

## table
filtered_data %>%
  group_by(measure) %>%
  summarise(Mean = mean(value, na.rm = TRUE) %>% round(digits = 1),
            'Standard deviation' = sd(value, na.rm = TRUE) %>% round(digits = 1)) %>% 
knitr::kable()

```
#### Continuous Value function
```{r vf Vertical Structure - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Vertical Structure - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments

VFs:  
- good general agreement
- may re-consider 2 and 3 layers value 

Weights:  
- most think middling importance
- a few think amongst most important, a few least important


## Native tree an shrub species richness 
### {.tabset .tabset-fade .tabset-pills}


```{r vf Native tree an shrub species richness - init, echo = F , results="asis"}

ind.num = 4
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Native tree an shrub species richness - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Categorised value function
```{r vf Native tree an shrub species richness - categorical, echo = F , results="asis"}
categorised_vf_fig()

## table
filtered_data %>%
  group_by(measure) %>%
  summarise(Mean = mean(value, na.rm = TRUE) %>% round(digits = 1),
            'Standard deviation' = sd(value, na.rm = TRUE) %>% round(digits = 1)) %>% 
knitr::kable()

```

#### Weights
```{r vf Native tree an shrub species richness - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments

VFs:  
- Feeling consensus may strengthen easily here
- Particular focus useful for 1-2 species  

Weightings - big variation, some making it least important, some making it most.  

## Invasives 
### {.tabset .tabset-fade .tabset-pills}


```{r vf Invasives - init, echo = F , results="asis"}

ind.num = 5
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Invasives - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Invasives - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments
VF:  
- a lot of variation. Worth considering Woodland Wildlife Toolkit approaches when revising estimates.

Weightings:  
- some think it mos important, rest middling.  

## Deadwood 
### {.tabset .tabset-fade .tabset-pills}

```{r vf Deadwood - init, echo = F , results="asis"}

ind.num = 6
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Deadwood - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Deadwood - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments

VFs:  
- A lot of variation, consensus may emerge

Weights:  
- Some think it most important (and are quite certain)  
- Some of lower or middling importance (but typically less certain

## AVTs 
### {.tabset .tabset-fade .tabset-pills}

```{r vf AVTs - init, echo = F , results="asis"}

ind.num = 7
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf AVTs - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf AVTs - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments

VFs:  
- Biggest disagreement is about where max should occur  

Weights:
- Massive variation, may need some evidence or guidance.  


## Woodland Extent 
### {.tabset .tabset-fade .tabset-pills}

```{r vf Woodland Extent - init, echo = F , results="asis"}

ind.num = 8
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Woodland Extent - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Woodland Extent - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments

VFs:  
- consensus may form  

Weights:  
- Some think it most important, some more middling


## Regeneration 
### {.tabset .tabset-fade .tabset-pills}

```{r vf Regeneration - init, echo = F , results="asis"}

ind.num = 9
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```  
  
#### Continuous Value function
```{r vf Regeneration - continuous, echo = F , results="asis"}
continuous_vf_fig()
```


#### Categorised value function
```{r vf Regeneration - categorical, echo = F , results="asis"}
categorised_vf_fig()

## table
filtered_data %>%
  group_by(measure) %>%
  summarise(Mean = mean(value, na.rm = TRUE) %>% round(digits = 1),
            'Standard deviation' = sd(value, na.rm = TRUE) %>% round(digits = 1)) %>% 
knitr::kable()

```
#### Weights
```{r vf Regeneration - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments

VFs:  
- Disagreement in shape of relationship
- A lot of variation around value of 1 and 2 classes

Weights:  
- a lot of variation

## Herbivore Impact 

Note: Complete revision of indicator since ROUND 1

### {.tabset .tabset-fade .tabset-pills}

```{r vf Herbivore Impact - init, echo = F , results="asis"}

ind.num = 10
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Herbivore Impact - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Herbivore Impact - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments

## Tree Health 
### {.tabset .tabset-fade .tabset-pills}

```{r vf Tree Health - init, echo = F , results="asis"}

ind.num = 11
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Tree Health - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Tree Health - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments

VFs:  
- A lot of variation - Refer to Woodland Wildlife Toolkit approach when reviewing

Weights:  
- full range of variation

## Ground Flora
### {.tabset .tabset-fade .tabset-pills}

```{r vf Ground Flora - init, echo = F , results="asis"}

ind.num = 12
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Ground Flora - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Ground Flora - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments  

Will obviously need development, particularly RE appropriate species for the site.  

VF:  
- logarithmic "r" shaped relationship most popular, followed by sigmoid "s" shape
- Consensus likely

weights: medium - high importance


## Horizontal complexity 
### {.tabset .tabset-fade .tabset-pills}

```{r vf Horizontal complexity - init, echo = F , results="asis"}

ind.num = 13
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Horizontal complexity - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Categorised value function
```{r vf Horizontal complexity - categorical, echo = F , results="asis"}
categorised_vf_fig()

## table
filtered_data %>%
  group_by(measure) %>%
  summarise(Mean = mean(value, na.rm = TRUE) %>% round(digits = 1),
            'Standard deviation' = sd(value, na.rm = TRUE) %>% round(digits = 1)) %>% 
knitr::kable()

```
#### Weights
```{r vf Horizontal complexity - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)
```

### Comments  

VFs:  
- consensus likely

Weights: big range

## Human Damage 
### {.tabset .tabset-fade .tabset-pills}

```{r vf Human Damage - init, echo = F , results="asis"}

ind.num = 14
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]
indicator_name  

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

paste(length(unique(filtered_data$respondant_name)), "respondants.")
```

#### Continuous Value function
```{r vf Human Damage - continuous, echo = F , results="asis"}
continuous_vf_fig()
```

#### Weights
```{r vf Human Damage - weights, fig.width= 4, echo = F , results="asis"}

ggplot_resp_weight_ind(indicator_name = indicator_name)

```

### Comments  

VFs:
- Big range, but may converge to a consensus.  
- Recommend using WWT method to contextualize

Weights:  
- Full range

# Certainties



# Discussion points
- [ ] - Some value functions have a minimum score >0. Strictly speaking this is okay... but it does meant that the final condition score will not be on a 0-100 scale (there will always be some minimum condition score).







