---
title: "Exploring Old condition data"
---

Prior to the WCA project, data was collected to inform management on WT estate sites, with condition inferred mainly from the subjective assessment of whether or not management interventions were desirable or practical for the different indicators measured. Notably, the survey employed no structured sampling design, and took a very different approach to the measurement of indicators.

With that in mind, a bowel of salt should be liberally added to any interpretation of the results of any analysis considering or comparing data collected from them with that from the new approach.

### Caveat: a very different approach to indicator estimation and scoring

Using the old method, indicators were estimated at the survey-zone (or "strata") level, rather than within sample plots[^exploring-old-condition-data-1], using a relatively subjective method. Estimating at zone-scale means that the size of the zone can have a very big impact on indicator measurements:

[^exploring-old-condition-data-1]: Plots were used in the old method but they were optional and not a structured sample. Further, the measurements taken were not averaged up to the zone level formally. Rather, scaling up to the zone was intuition-based, using the plots as a guide, along with the proportion areas of the woodland that they survey thought they represented.

-   Larger zones will likely have more species

-   Large areas are expected to contain more variation (large size captures more unusual obervations), but this is likely to be dampened in estimation due to averaging over those larger areas[^exploring-old-condition-data-2].

[^exploring-old-condition-data-2]: Known as "regression towards the mean"

The survey's (optional) plot-stops collected detailed information (though, not their location). The Sylva/FC method also typically uses zone-average measurements for many of the indicators.

> *We are not assessing the condition of an imagined "average" area of woodland... We are assessing the condition of actual locations that represent the wider woodland*

Our WECA approach typically measures and scores[^exploring-old-condition-data-3] indicators at plots, and then averages those scores across the survey-zone. This difference is important (averaging plot-level indicator value scores, rather than averaging their measurements then scoring that zone-wide average). WECA aims to represent the ecological condition of actual real-life locations within a wood, not the condition of an imagined area of "average" woodland. The thought experiment below illustrates the potential implications of these different approaches for scoring one indicator.

[^exploring-old-condition-data-3]: using a value function

![](){fig-align="center"}

```{r libraries}
library(tidyverse)
```

```{r load data}

cnd <- read.csv("Data//old condition data//Condition Assessment Data 2025_01.csv")

names(cnd)



```

```{r ininital-curation}
missing_vals <- c("NULL", "null", " ", "N/A", "", "NA", "n/a", "Null")

#cbind(cnd$OpenSpaceRidesPercentMax.1, cnd$OpenSpaceRidesPercentMax) # dupliocated col, needs removed
cnd <- cnd %>%
  select(-"OpenSpaceRidesPercentMax.1") %>% # removed duplicated col
# rename cols RTS.Intervention.Desirable.1 RTS.Intervention.Achievable.1
names(cnd[,c("RTS.Intervention.Desirable.1" , "RTS.Intervention.Achievable.1")]) <-   c("RSS.Intervention.Desirable" , "RSS.Intervention.Achievable")


cnd <- cnd %>%
  mutate(across(where(is.character),
                ~iconv(.x, from = "", to = "UTF-8", sub = " "))) %>% # 
  mutate(across(everything(), ~{
    x <- trimws(as.character(.))

    # Standardize missing values
    x[x %in% c("", " ", "NULL", "null", "NA", "N/A", "n/a", "Null")] <- NA

    # yes/no → TRUE/FALSE
    x[tolower(x) == "yes"] <- TRUE
    x[tolower(x) == "no"]  <- FALSE

    # If column now only contains logical-like values, convert it
    if (all(x %in% c("TRUE", "FALSE", NA))) x <- as.logical(x)

    x
  }))





# format to have two data frames, one for the event of the survey, the other for data collected, both with the id coloum to allow matching

# survey_info <-  
## coloums for "id", "CostCentre", "Site.Name", "SiteManager", "Region", "PlannedObservationDate", "ActualObservationDate", "Surveyor", "Whole.Site", "Key.Feature", "Stratum", "Stratum.Description", "Overall.Intervention.Desirable", "Overall.Intervention.Achievable", "Overall.Conclusions.And.Recommendations", "TotalTreeCanopyCoverPercentCurrent", "TotalTreeCanopyCoverPercentMin", "TotalTreeCanopyCoverPercentMax", "OpenSpaceSemiNaturalHabitatPercentCurrent", "OpenSpaceSemiNaturalHabitatPercentMin", "OpenSpaceSemiNaturalHabitatPercentMax", "OpenSpaceRidesPercentCurrent", "OpenSpaceRidesPercentMin", "OpenSpaceRidesPercentMax", "OpenSpaceTemporaryPercentCurrent", "OpenSpaceTemporaryPercentMin", "OpenSpaceWaterFeaturesPercentCurrent", "OpenSpaceWaterFeaturesPercentMin", "OpenSpaceWaterFeaturesPercentMax", "ShrubCoverPercentCurrent", "ShrubCoverPercentMin", "ShrubCoverPercentMax", "Intervention.Desirable", "Intervention.Achievable", "NotesAction", "TAIntervention.Desirable", "TA.Intervention.Achievable", "TA.NotesActions", "TS.Intervention.Desirable", "TS.Intervention.Achievable", "TS.NotesActions", "SS.Intervention.Desirable", "SS.Intervention.Achievable", "SS.NotesActions", "LTR.Intervention.Desirable", "LTR.Intervention.Achievable", "LTR.NotesActions", "LSR.Intervention.Desirable", "LSR.Intervention.Achievable", "LSR.NotesActions", "RTS.Intervention.Desirable", "RTS.Intervention.Achievable", "RTS.NotesActions", "RSS.Intervention.Desirable" , "RSS.Intervention.Achievable", "RSS.NotesActions", "Flora.Intervention.Desirable", "Flora.Intervention.Achievable", "Flora.NotesActions", "Deadwood.Intervention.Desirable", "Deadwood.Intervention.Achievable", "Deadwood.NotesActions", "Animal.Damage.Intervention.Desirable", "Animal.Damage.Intervention.Achievable", "Animal.Damage.NotesActions", "Invasives.Intervention.Desirable", "Invasives.Intervention.Achievable", "Invasives.NotesActions", "Tree.Health.Intervention.Desirable", "Tree.Health.Intervention.Achievable", "Tree.Health.NotesActions", "Human.Impacts.Intervention.Desirable", "Human.Impacts.Intervention.Achievable", "Human.Impact.Notes"

## rename "Intervention.Desirable" and "Intervention.Achievable" "TS_cover_intervention..."
## add a col for N plots based on max plot number



# old_condition_data <-  
## cols: "id", PLOT NUMBER, indicator name, measure



```
