---
title: "Exploring Old condition data"
---

Prior to the WCA project, data was collected to inform management on WT estate sites, with condition inferred mainly from the subjective assessment of whether or not management interventions were desirable or practical for the different indicators measured. Notably, the survey employed no structured sampling design, and took a very different approach to the measurement of indicators.

With that in mind, a bowel of salt should be liberally added to any interpretation of the results of any analysis considering or comparing data collected from them with that from the new approach.

### Caveat: Inconsistent data collection

The data using the old method often includes blank or NULL values. Often, it is clear that this is inferred as an absence of the feature being measured (e.g. no deadwood present), here analysis assumes that - where data exists for similar indicators within a plot - that missing data implies a "0", "absence" or "FALSE" for the indicator, as appropriate. 

It should also be noted, that there are a number of improbable observations within the data: e.g. where a tree/shrub species cover is present, but the species count is 0. These instances have typically been left as they are.


### Caveat: a very different approach to indicator estimation and scoring

Using the old method, indicators were estimated at the survey-zone (or "strata") level, rather than within sample plots[^exploring-old-condition-data-1], using a relatively subjective method. Estimating at zone-scale means that the size of the zone can have a very big impact on indicator measurements:

[^exploring-old-condition-data-1]: Plots were used in the old method but they were optional and not a structured sample. Further, the measurements taken were not averaged up to the zone level formally. Rather, scaling up to the zone was intuition-based, using the plots as a guide, along with the proportion areas of the woodland that they survey thought they represented.

-   Larger zones will likely have more species

-   Large areas are expected to contain more variation (large size captures more unusual obervations), but this is likely to be dampened in estimation due to averaging over those larger areas[^exploring-old-condition-data-2].

[^exploring-old-condition-data-2]: Known as "regression towards the mean"

The survey's (optional) plot-stops collected detailed information (though, not their location). The Sylva/FC method also typically uses zone-average measurements for many of the indicators.

> *We are not assessing the condition of an imagined "average" area of woodland... We are assessing the condition of actual locations that represent the wider woodland*

Our WECA approach typically measures and scores[^exploring-old-condition-data-3] indicators at plots, and then averages those scores across the survey-zone. This difference is important (averaging plot-level indicator value scores, rather than averaging their measurements then scoring that zone-wide average). WECA aims to represent the ecological condition of actual real-life locations within a wood, not the condition of an imagined area of "average" woodland. The thought experiment below illustrates the potential implications of these different approaches for scoring one indicator.

[^exploring-old-condition-data-3]: using a value function

![](){fig-align="center"}

```{r libraries}
library(tidyverse)
```
```{r constants}
# Regex patterns to detect plot numbers and overall columns
plot_regex <- "(\\.|_)?[Pp]lot\\.?([0-9]+)$"  # matches variants like Plot1, plot.2, _plot3
overall_regex <- "\\.Overall$"                # matches columns ending in .Overall

```
```{r cnd_load}
cnd <- read.csv("Data//old condition data//Condition Assessment Data 2025_01.csv")

```
```{r cnd_rename_cols}

curation_notes <- character()
results <- list()

#cbind(cnd$OpenSpaceRidesPercentMax.1, cnd$OpenSpaceRidesPercentMax) # dupliocated col, needs removed
cnd <- cnd %>%
  select(-"OpenSpaceRidesPercentMax.1") %>% # removed duplicated col
  # Rename "Intervention.Desirable" and "Intervention.Achievable" to be explicit RE cover of tree species, shrub species and open space of different types
  rename(
    Tree.shrub.space.cover_Intervention.Desirable = Intervention.Desirable,
    Tree.shrub.space.cover_Intervention.Achievable = Intervention.Achievable,
    # rename cols RTS.Intervention.Desirable.1 RTS.Intervention.Achievable.1
    RSS.Intervention.Desirable = RTS.Intervention.Desirable.1,
    RSS.Intervention.Achievable = RTS.Intervention.Achievable.1
  )

# standardising plot/overall indicator nomeculature in col names ----
# Function to canonicalise a column name while keeping dots and adding Plot/Overall suffix
canonicalise_colname <- function(x) {

  # 1️⃣ Extract plot number (if any)
  plot_num <- str_extract(x, plot_regex)
  plot_num <- ifelse(is.na(plot_num), NA, str_extract(plot_num, "[0-9]+"))

  # 2️⃣ Detect Overall
  is_overall <- grepl(overall_regex, x)

  # 3️⃣ Strip Plot / Overall FIRST
  core <- x %>%
    gsub(plot_regex, "", ., ignore.case = TRUE) %>%   # remove Plot suffix
    gsub(overall_regex, "", ., ignore.case = TRUE)    # remove Overall suffix

  # 4️⃣ Canonicalise indicator naming on the *core*
  core <- core %>%
    gsub("NativesNumberPresent$", ".Native.richness", ., ignore.case = FALSE) %>%
    gsub("NonNativesNumberPresent$", ".Non.Native.richness", ., ignore.case = FALSE) %>%
    gsub("\\.Natives\\.Present$", ".Native.richness", ., ignore.case = FALSE) %>%
    gsub("\\.NonNatives\\.Present$", ".Non.Native.richness", ., ignore.case = FALSE) %>% 
    gsub("\\SSDAFOROfCover$", "Shrub.Cover", ., ignore.case = FALSE) %>% 
      # Fix cases like: 
          # RSSNon.Native.richness  -> RSS.Non.Native.richness
          # RSS.Non.Native.richness -> unchanged
    gsub("^([A-Z]{2,4})\\.?Non\\.", "\\1.Non.", .)


  # 5️⃣ Reattach Plot / Overall in a consistent way
  if (!is.na(plot_num)) {
    paste0(core, "_Plot", plot_num)
  } else if (is_overall) {
    paste0(core, "_Overall")
  } else {
    core
  }
}

old.names <- names(cnd)
names(cnd) <- names(cnd) %>% sapply(canonicalise_colname) %>% as.character()

# some more renaming
cnd <- cnd %>%
  rename(
    # Rename "Tree.shrub.space.cover_Intervention.Desirable" and "Tree.shrub.space.cover_Intervention.Achievable" back
    SS.Native.richness_Overall = SSNativesPresent_Overall,
    SS.Non.Native.richness_Overall = SSNonNativesPresent_Overall,
    TS.Native.richness_Overall = TSNativesPresent_Overall,
    TS.Non.Native.richness_Overall = TSNonNativesPresent_Overall
  )

```
```{r cnd_standardise_missing_yesno}
missing_vals <- c("NULL", "null", " ", "N/A", "", "NA", "n/a", "Null", "N")
cnd <- cnd %>%
  mutate(across(where(is.character),
                ~iconv(.x, from = "", to = "UTF-8", sub = " "))) %>% # 
  mutate(across(everything(), ~{
    x <- trimws(as.character(.))

    # Standardize missing values
    x[x %in% missing_vals] <- NA

    # yes/no → TRUE/FALSE
    x[tolower(x) == "yes"] <- TRUE
    x[tolower(x) == "no"]  <- FALSE

    # If column now only contains logical-like values, convert it
    if (all(x %in% c("TRUE", "FALSE", NA))) x <- as.logical(x)

    x
  }))

curation_notes <- c(
  curation_notes,
  "- Standardised missing values (NULL, N/A, etc.) were converted yes/no to TRUE/FALSE"
  )
```
``` {r indicator_types_format_theme}

## ID indicator names  ----
indicator_names <- names(cnd) %>%
  grep("_Overall$", ., value = TRUE) %>%
  sub("_Overall$", "", .)

## check entries at each indicator ---- 

indicator_freq_tables <- indicator_names %>%
  setNames(., .) %>%   # name the list by indicator
  lapply(function(ind) {

    # all columns for this indicator (plots + overall)
    cols <- names(cnd) %>%
      grep(paste0("^", ind, "_"), ., value = TRUE)

    # combine values across plots + overall
    values <- cnd %>%
      dplyr::select(dplyr::all_of(cols)) %>%
      unlist(use.names = FALSE)

    # frequency table (including NA)
    table(values, useNA = "ifany")
  })

# initial check that no wild unexpected entries
# indicator_freq_tables 

## format indicator types ----

dafor_levels <- c("D", "A", "F", "O", "R", "Absent") %>% rev()# including "Absent" for missing data assumed as absence

is_numeric_chr <- function(x) {
  all(grepl("^[-+]?[0-9]*\\.?[0-9]+$", x))
}

is_logical_chr <- function(x) {
  all(x %in% c("TRUE", "FALSE"))
}

is_dafor_chr <- function(x, dafor_levels) {
  sum(x %in% dafor_levels) >= 3
}


indicator_types <- indicator_freq_tables %>%
  imap_dfr(function(tbl, indicator) {

    vals <- names(tbl)
    vals <- vals[!is.na(vals) & vals != "<NA>"]

    tibble(
      indicator    = indicator,
      n_unique     = length(vals),
      is_dafor     = is_dafor_chr(vals, dafor_levels),
      is_logical   = is_logical_chr(vals),
      is_numeric   = is_numeric_chr(vals)
    )
  }) %>%
  mutate(
    type = case_when(
      is_dafor   ~ "dafor",
      is_logical ~ "logical",
      is_numeric ~ "numeric",
      TRUE       ~ "unknown"
    )
  ) 

# group indicators by theme -----
indicator_types$theme <- NA_character_
indicator_types$theme[grep("^TA\\.", indicator_types$indicator)] <- "tree_age"
indicator_types$theme[grep("^(TS|SS)\\.", indicator_types$indicator)] <- "tree_shrub_species_richness"
indicator_types$theme[indicator_types$indicator %in% c("TS.Canopy.Dominated.By.One.Or.Two.SPP",
                                                       "SS.Shrub.Layer.Dominated.By.One.Or.Two.SPP") ] <- "tree_shrub_species_dominated"

indicator_types$theme[grep("Shrub.Cover", indicator_types$indicator)] <- "shrub_cover"
indicator_types$theme[grep("^(LTR|LSR)\\.", indicator_types$indicator)] <- "regeneration_level"
indicator_types$theme[grep("^(RTS|RSS)\\.", indicator_types$indicator)] <- "regeneration_species_richness"
indicator_types$theme[indicator_types$indicator %in% c("RTS.Dominated.By.One.Or.Two.SPP",
                                                       "RSS.Dominated.By.One.Or.Two.SPP") ] <- "regeneration_species_dominated"

indicator_types$theme[grep("^Flora\\.", indicator_types$indicator)] <- "flora"
indicator_types$theme[grep("^Deadwood\\.", indicator_types$indicator)] <- "deadwood"
indicator_types$theme[grep("^Invasives\\.", indicator_types$indicator)] <- "invasives"
indicator_types$theme[grep("^Animal\\.Damage\\.", indicator_types$indicator)] <- "animal_damage"
indicator_types$theme[grep("^Human\\.Impacts\\.", indicator_types$indicator)] <- "human_impact"
indicator_types$theme[grep("^Tree\\.Health\\.", indicator_types$indicator)] <- "tree_health"

indicator_types <- indicator_types %>% 
  arrange(theme)



```
```{r f_format_indicators}

# function to get all cols for an indicator ----
indicator_cols <- function(df, indicator) {
  grep(
    paste0("^", indicator, "_(?:Plot[0-9]+|Overall)$"),
    names(df),
    value = TRUE
  )
}

# function to format indicators ----
format_indicators <- function(df, indicator_types) {


  for (i in seq_len(nrow(indicator_types))) {
    ind  <- indicator_types$indicator[i]
    type <- indicator_types$type[i]
    cols <- indicator_cols(df, ind)

    if (length(cols) == 0) next
    if (type == "dafor") {
      df <- df %>%
        mutate(
          across(
            all_of(cols),
            ~ factor(.x, levels = dafor_levels, ordered = TRUE)
          )
        )} else if (type == "logical") {
          df <- df %>%
            mutate(across(all_of(cols), as.logical)
                   )
          } else if (type %in% c("numeric", "richness")) {
            df <- df %>%
              mutate(across(all_of(cols), as.numeric))
          }}
  df
}

```
```{r cnd_type_formatting}
# Apply type formatting to cnd ----
cnd <- format_indicators(cnd, indicator_types)

curation_notes <- c(
  curation_notes,
  "- Indicator column names were standardised to have consistent 'Plot'/'Overall' suffixes"
  )


```
```{r plot-level-data-extraction}
# plot-level indicator measurements ----
plot_cols <- names(cnd)  %>% 
  grep(plot_regex, ., value = TRUE, ignore.case = TRUE)

## want a survey tidy data frame with:
### id, Site.Name, ActualObservationDate, Key.Feature, Stratum, Plot.Number, measure for each indicator

# initial curation ----
long_plots  <- cnd %>%
  select(
    id,
    Site.Name,
    ActualObservationDate,
    Key.Feature,
    Stratum,
    all_of(plot_cols)
  ) %>%
  # pivot longer to get col for indicator and plot number
  mutate(across(all_of(plot_cols), as.character)) %>% ## covert to character to ensure pivot works
  pivot_longer( # lengthen to get indicator and plot number cols
    cols = all_of(plot_cols),
    names_to = "indicator_plot",
    values_to = "value"
  ) %>%
  mutate( # use to create Plot.Number and indicator cols
    Plot.Number = str_extract(indicator_plot, "(?<=_Plot)\\d+") %>% as.integer(),
    indicator   = str_remove(indicator_plot, "_Plot\\d+$")
  ) %>%
  # select relevant cols to take forward
  select(
    id,
    Site.Name,
    ActualObservationDate,
    Key.Feature,
    Stratum,
    Plot.Number,
    indicator,
    value
  ) %>%
  # widen to have one col per indicator
  pivot_wider(
    names_from  = indicator,
    values_from = value
  ) %>%
  # drop all rows with NA throuhgout all cols with names in indicator_types$indicator
  filter(
    if_any(
      all_of(indicator_types$indicator),
      ~ !is.na(.x)
    )
  ) %>% 
  arrange(id, Plot.Number) %>% 
  as.data.frame() %>% 
  format_indicators(., indicator_types)


curation_notes <- c(
  curation_notes,
  "- Plot-level data was extracted and converted into long format data frame (row for each plot)"
  )

results$n.plots_og <- nrow(long_plots)  
results$n.plot_sites_og <- length(unique(long_plots$Site.Name))
```
```{r plot_level_formatting}
# formating indicator types as in cnd ----
## Apply type formatting

# format each indicator coloum data type ----
for (i in seq_along(indicator_types$indicator)) {
  ind <- indicator_types$indicator[i]
  
  # look up the type for this indicator
  ind_type <- indicator_types$type[indicator_types$indicator == ind]
  
  # apply formatting to the entire column
  if (ind_type == "dafor") {
    long_plots[, ind] <- factor(
      long_plots[, ind],
      levels  = dafor_levels,   # e.g. c("A","D","F","O","R")
      ordered = TRUE
    )} else if (ind_type == "logical") {
      long_plots[, ind] <- as.logical(long_plots[, ind])
      } else if (ind_type %in% c("numeric", "richness")) {
        long_plots[, ind] <- as.numeric(long_plots[, ind])
        } else {
          # unknown type → leave unchanged
          long_plots[, ind] <- long_plots[, ind]
        }
}
```
```{r plot_data_data_manipulation}
curation_notes <- c(
  curation_notes,
  paste0( "- The initial plot data contained information from ", results$data_description$nplots_og, " plots, covering ", results$data_description$n_surveys_og, " surveys over ",  results$data_description$nplot_sites_og, " sites" )
)

source("Scripts//old_condition_plot_data_curation.R")

```
```{r create_survey_level_info}

# survey_info <-  all relevent info collected at survey-level scale  ----

## explain ----
## coloums for:
### survey info - "id", "CostCentre", "Site.Name", "SiteManager", "Region", "PlannedObservationDate", "ActualObservationDate", "Surveyor", "Whole.Site", "Key.Feature", "Stratum", "Stratum.Description", 
### cover - "TotalTreeCanopyCoverPercentCurrent", "ShrubCoverPercentCurrent",  "OpenSpaceSemiNaturalHabitatPercentCurrent", "OpenSpaceRidesPercentCurrent", "OpenSpaceTemporaryPercentCurrent", "OpenSpaceWaterFeaturesPercentCurrent", 
### target covers - "TotalTreeCanopyCoverPercentMin", "TotalTreeCanopyCoverPercentMax", "ShrubCoverPercentMin", "ShrubCoverPercentMax",  "OpenSpaceSemiNaturalHabitatPercentMin", "OpenSpaceSemiNaturalHabitatPercentMax", "OpenSpaceRidesPercentMin", "OpenSpaceRidesPercentMax","OpenSpaceTemporaryPercentMin", "OpenSpaceWaterFeaturesPercentMin", "OpenSpaceWaterFeaturesPercentMax", 
### intervensions, reccomendationsm notes and actions - "Overall.Intervention.Desirable", "Overall.Intervention.Achievable", "Overall.Conclusions.And.Recommendations",      "Intervention.Desirable", "Intervention.Achievable", "NotesAction", "TAIntervention.Desirable", "TA.Intervention.Achievable", "TA.NotesActions", "TS.Intervention.Desirable", "TS.Intervention.Achievable", "TS.NotesActions", "SS.Intervention.Desirable", "SS.Intervention.Achievable", "SS.NotesActions", "LTR.Intervention.Desirable", "LTR.Intervention.Achievable", "LTR.NotesActions", "LSR.Intervention.Desirable", "LSR.Intervention.Achievable", "LSR.NotesActions", "RTS.Intervention.Desirable", "RTS.Intervention.Achievable", "RTS.NotesActions", "RSS.Intervention.Desirable" , "RSS.Intervention.Achievable", "RSS.NotesActions", "Flora.Intervention.Desirable", "Flora.Intervention.Achievable", "Flora.NotesActions", "Deadwood.Intervention.Desirable", "Deadwood.Intervention.Achievable", "Deadwood.NotesActions", "Animal.Damage.Intervention.Desirable", "Animal.Damage.Intervention.Achievable", "Animal.Damage.NotesActions", "Invasives.Intervention.Desirable", "Invasives.Intervention.Achievable", "Invasives.NotesActions", "Tree.Health.Intervention.Desirable", "Tree.Health.Intervention.Achievable", "Tree.Health.NotesActions", "Human.Impacts.Intervention.Desirable", "Human.Impacts.Intervention.Achievable", "Human.Impact.Notes"

# 1. Identify columns 
## Survey metadata columns
survey_metadata_cols <- c(
  "id","CostCentre","Site.Name","SiteManager","Region",
  "PlannedObservationDate","ActualObservationDate","Surveyor",
  "Whole.Site","Key.Feature","Stratum","Stratum.Description"
)
## intervention/notes 
intervention_cols <- grep("Intervention|NotesAction|NotesActions|Recommendations", 
                           names(cnd), value = TRUE)
## cover (current) and target (min/max) columns
cover_current_cols <- grep("PercentCurrent$", names(cnd), value = TRUE)
cover_target_cols  <- grep("Percent(Min|Max)$", names(cnd), value = TRUE)
## indicators estimated at overall-level scale 
indicator_overall_cols <- grep("_Overall$", names(cnd), value = TRUE)

# 2. Build survey_info
survey_info <- cnd %>%
  select(all_of(survey_metadata_cols),
         all_of(cover_current_cols),
         all_of(cover_target_cols),
         all_of(intervention_cols),
         all_of(indicator_overall_cols))

# 3. Add number of plots
plot_nums <- str_extract(names(cnd), "_Plot\\d+$") %>%  # "_Plot1", "_Plot2", etc.
  str_remove("_Plot") %>%                   # remove "_Plot" prefix
  as.numeric()   
keep_cols <- which(!is.na(plot_nums)) # cols with plot info
plot_nums_kept <- plot_nums[keep_cols] %>% as.matrix()   # these match plot_ind_present columns
plot_ind_present <- !is.na(cnd[, keep_cols])*1
that_plot_nums <- sweep(plot_ind_present, 2, plot_nums_kept, `*`) %>% 
  replace(., . == 0, NA) # give everything its plot number where present, else NA
survey_info$max_plot <- that_plot_nums %>%
  # max of each row
  apply(1, function(x) if(all(is.na(x))) NA else max(x, na.rm = TRUE))  %>% 
  #replace NA with 0
  replace(., is.na(.), 0)  
rm(plot_ind_present)

### troubleshooting ----
# table(length(survey_info$max_plot))
# #### explore frequency of each number of plots
# df_freq <- as_tibble(that_plot_nums) %>%
#   mutate(row_id = row_number()) %>%
#   pivot_longer(-row_id, names_to = "col", values_to = "plot") %>%
#   filter(!is.na(plot) & plot != 0) %>%   # ignore 0 / NA
#   count(row_id, plot) %>%                # frequency table
#   pivot_wider(names_from = plot,
#               values_from = n,
#               values_fill = 0)  %>%          # fill missing with 0
#   
# ### looks like there are 4 indicators that area always present
# that_plot_nums[1, which(that_plot_nums[1, ]==4)]

```


Some curation decisions are documented below.

- Some columns indicating whether plots were dominated by one or two species had "no" values where it was believed data was not collected (i.e. other data was NA). These were converted to NA.
- The DAFOR scale has no "0" value. Absence was assumed when NA was entered in plots that had data recorded elsewhere for that indicator type, FALSE was assumed for logical data where data appeared similarly missing. 
