# Introduction {.unnumbered}

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
# library(readxl)
library(tidyverse)
# library(stringr)
# library(plotly)
library(htmltools)
library(htmlwidgets)
# library(gridExtra)
# library(grid)
library(DT)
# library(ggiraph)
# library(cowplot)
# library(ggnewscale)
library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

 forms.direct <- "Data\\Delphi round 1\\editing for round 2\\"
# forms.direct <- "Data\\Delphi round 1\\response sheets\\"

extraction.location <- "Data\\Delphi round 1\\"

```

```{r data_extraction_from_excelsheets and curation - preliminary work, echo = F, include = F, eval = F }

# extract from excel sheets ----
source("Scripts\\Extract_expert_info.R")
# if all is well this should:
# save ind.matcher ind.matcher.df and expert.data (expert opinion responses from the Delphi forms) 
# as files within the specified extraction.location to be loaded together

# curation ----
source("Scripts\\curation of extracted data.R")
# this one loads the data and does some curation, scaling responses etc
# then saves the dataframes

# then later I loads it (in Italian-American mafiosi accent)

```

```{r load curated data, echo = F, include = F, eval = T }
load(paste0(extraction.location,"curated.RData"))
```

A lot of the figures and tables in this book are interactive, you can often hover to get more info, click to filter/sort different data etc. Please get in touch if any sections need clarification, if you notice any bugs or errors, or have any other feedback that might improve the doc ([EwanMcHenry\@woodlandtrust.org.uk](mailto:EwanMchenry@woodlandtrust.org.uk)).

The book explores `r has_completed$indicator_name %>% unique() %>% length()` indicators of woodland ecological condition: how they can be measured and used to assess the ecological condition of woodland sites, based on a structured synthesis of evidence from the literature and the expert opinion of professional woodland conservationists.

The proposed framework takes inspiration from many existing projects internationally that inform management and policy through an assessment of ecological condition at different scales, measured using structural indicators.

## A Definition

> **Woodland ecological condition (WEC)**: The potential for a woodland site - in its current structural state - to sustain biodiverse, dynamic and well-functioning species communities (i.e. ignoring time-lags in wildlife species colonisation).

## Book structure

1.  [Introduction (this page)](index.qmd)

2.  [Instructions to experts for round 2 of Delphi process](002_delphi_r2_instructions.qmd)

3.  [Project background and scope](001_project%20background.qmd)

4.  [An explanation of the general approach to estimate condition](01_general%20approach.qmd)

5.  A series of pages for each indicator in turn exploring:

    -   Justification of its inclusion
    -   A proposed measurement method
    -   An exploration of value functions and weightings
        -   Tabs can be used to switch view of value functions from continuous lines to classes (e.g. verticle structure, where only 5 options are possible), and explore importance weightings (compared to other indicators).
    -   Further considerations for indicator development
    -   Exceptions where a general approach may require adaptation.
        -   Indicatively categorised as "priority", "lower priority", and exceptions dealt with by already/with adaption/by other indicators. Some reasoning is typically given for categorisation.
    -   Approaches to this indicator from other organisations

6.  An exploration of the [importance weightings](04_Importance%20weightings.qmd) for each indicator

7.  [The methods to gather and analyse evidence, particularly expert opinion](02_methods.qmd)
