---
title: "Assessing Woodland Ecological Condition"
date:    |
  `r format(Sys.time(), '%d %B, %Y')`
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
---
<style>
.html-widget {
  max-width: 1000px;
  margin: auto;
}
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
forms.direct <- "Delphi round 1\\response sheets\\"
extraction.location <- "Delphi round 1\\"

```

```{r data_extraction_from_excelsheets - preliminary work, echo = F, include = F, eval = F }

# ---------------------------------------------------------------
# RUN SCRIPT TO EXTACT FROM EXCEL SHEETS ----------------------------------
# ---------------------------------------------------------------
source("Extract_expert_info.R")
# if all is well this should:
# save ind.matcher ind.matcher.df and expert.data (expert opinion responses from the Delphi forms) 
# as files within the specified extraction.location to be loaded together
# then later I loads it (in Italian-American mafiosi accent)
```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
library(readxl)
library(tidyverse)
library(stringr)
library(plotly)
library(ggpubr)
library(mgcv)
library(RColorBrewer)
library(htmltools)
library(htmlwidgets)
library(gridExtra)
library(grid)
library(DT)
library(Polychrome)


#-----------------------------------------------------
#  LOAD DATA ----------------------------------------------

load(paste0(extraction.location,"sheets_data.RData"))

# ---------------------------------------------------------------
# ok, that was "fun", lets get to woooork!

```
```{r data_curation, echo = F, include = F, eval = T }
#-----------------------------------------------------
# DATA CURATION --------------------------------------------------
#-----------------------------------------------------
## Create a data frame of respondant data ------------
# Step 1: Extract data from expert.data list
respondent_names <- expert.data$respondant_name %>% unlist
indicator_nums <- expert.data$indicator_num %>% unlist
indicator_names <- expert.data$indicator_name %>% unlist
cert_val_funct <- expert.data$value_confidence_score %>% unlist
measure <- unlist(lapply(expert.data$value.points, `[[`, "measure"))
value <- unlist(lapply(expert.data$value.points, `[[`, "value"))
weight <- expert.data$weight_indicy_score %>% unlist()
cert_weight <- expert.data$weight_indicy_confidence %>% unlist()
vf.sentance <- expert.data$descriptive_sentance %>% unlist()

# Create a data frame by combining the data, scale and add variables
# 
df <- data.frame(
  respondant_name = rep(respondent_names, sapply(expert.data$value.points, nrow)),
  indicator_num = rep(indicator_nums, sapply(expert.data$value.points, nrow)),
  indicator_name = rep(indicator_names, sapply(expert.data$value.points, nrow)),
  
  measure = measure,
  value = value,
  cert_val_funct = rep(cert_val_funct, sapply(expert.data$value.points, nrow)),
  vf.sentance = rep(vf.sentance, sapply(expert.data$value.points, nrow)),
  
  weight = rep(weight, sapply(expert.data$value.points, nrow)),
  cert_weight = rep(cert_weight, sapply(expert.data$value.points, nrow)),
  
  stringsAsFactors = FALSE
) %>% merge(., ind.matcher.df, by.x = "indicator_name", by.y = "indicator_name", all.x = T) # this is wrong -- its sheet name.. but its fine


## scale and augment ----

df.scaled <- df %>% 
  group_by(indicator_name, respondant_name) %>% 
  mutate(value = ((value - min(value)) * 100)/ (max(value) - min(value)) # scale by min and max value for each respondant 
        ) 
df.scaled <- df.scaled %>% 
  group_by(respondant_name) %>% 
  mutate(weight = weight * (100 / max(weight)) # scale by biggest value for each respondant 
  )

which.scaled.VF <- paste(df$respondant_name, "-", df$indicator_name)[df.scaled$value != df$value] %>% 
  sort() %>%   unique()
which.scaled.wt <- paste(df$respondant_name)[df.scaled$weight != df$weight] %>% 
  sort() %>%   unique()

df$value <- df.scaled$value %>% round(0)
df$weight <- df.scaled$weight %>% round(0)

df$value.dec = df$value/100 # a decimal version of value to work in the later binomial gams

## version with sinlge entry for each indicator:respondant ----
just.one.df <- df[!duplicated(paste0(df$sheet_name, df$respondant_name)),
c("indicator_name", "respondant_name", "cert_val_funct", "cert_weight", "sheet_name", "weight")]

just.one.df$mean_repond_cert_vf <- ave(just.one.df$cert_val_funct, just.one.df$respondant_name, FUN = function(x) mean(x, na.rm = TRUE))
just.one.df$mean_repond_cert_wt <- ave(just.one.df$cert_weight, just.one.df$respondant_name, FUN = function(x) mean(x, na.rm = TRUE))
#-----------------------------------------------------

```
```{r participation_data_curation, echo = F, include = F, eval = T }
# completed df
has_completed <- data.frame(respondent_names = respondent_names,
                            indicator_name = indicator_names,
                            wt = !is.na(weight),
                            wt_cert = !is.na(cert_weight),
                            vf = sapply(expert.data$value.points, nrow)>1,
                            vf_cert = !is.na(as.numeric(cert_val_funct))
                            ) %>% 
  left_join(., ind.matcher.df[, c("sheet_name", "indicator_name")], by = "indicator_name")

# create df of group by(indicator, respondant) %>% sum(completed) for each
completed_summary <- has_completed %>% 
  group_by(respondent_names) %>% 
  summarize(vf = sum(vf),
            vf_cert = sum(vf_cert),
            wt = sum(wt),
            wt_cert = sum(wt_cert))
```


This document explores the results of expert opinion gathered from staff at the Woodland Trust, related to the measurement of woodland ecological condition by a set of `r has_completed$indicator_name %>% unique() %>% length()` indicators. 

# Methods
## Expert opinion elicitation using the Delphi method.  

Here, we estimate the value function and weight for each indicator using a method of expert opinion elicitation called the Delphi method. The Delphi method is designed to collect and distill expert knowledge, using iterative or repetitive surveys to increase robustness 
<!-- (Edwards et al., 2011) -->
. 
The method is also useful for identifying where agreements and disagreements occur.
<!-- (Buckley, 1995; Kangas, Alho, Kolehmainen, & Mononen, 1998; Landeta, 2006). --> 
  
The method relies on repetitive surveying in at least two to three‘rounds’ where participants are asked the same questions, to which they provide quantitative answers. After each round participants review and consider the answers given by all panel-members, and are then given the opportunity to revise their answers. Results are considered relatively robust when they change little between rounds. The Delphi method has a history of successful use in supporting conservation action where empirical evidence is insufficient and information is needed rapidly to inform decisions. It's transparency and repeatability also add to its apeal.

## Survey questions

Each participant either attended an introductory workshop to the process, or was taken through individually. A sub-section of participants were also involved in development the proposed methods to measure each indicator in the field. Following this participants were supplied with an Excel spread sheet in which to submit their answers to questions relating to:  

- Value functions describing how each indicator typically related to WEC
- Weightings decreeing the relative important of each indicator for WEC, typically
- Qualitative comments relating to: situations where these "typical" responses might be inappropriate (exceptions), concerns with the proposed method to measure the indicator, and any other information though relevant.

Participants were encouraged to keep a clear distinction between:  

- The quantitative responses they provided regarding the relationship between indicators and WEC in a typical woodland
- The exceptions to those typical relationships that they highlighted.

It was also stated that noted exceptions would be analysed and, where appropriate, used to inform future development of the WEC measure to account for important nuance.


## Indicators considered  

A set of `r has_completed$indicator_name %>% unique() %>% length()` structural indicators were proposed to measure WEC.

```{r  indicators table, echo = F }
ind.table <- ind.matcher.df [,1:2]
names(ind.table) <- c("Indicator", "Description")
ind.table %>% 
   datatable(extensions = c('FixedColumns',"FixedHeader"),
              options = list(scrollX = T, 
                             paging = F,
                             fixedHeader=F,
                             dom = 't',
                             fixedColumns = list(leftColumns = 1) #,
                             # fixedColumns = list(leftColumns = 1) ,
                             # buttons = c('copy', 'csv')
                             ),
              rownames = F#,
              #container = sketch.lcm.areas
              ) %>% 
    formatStyle( columns = c("Indicator"), fontWeight = 'bold', `text-align` = 'left', "white-space" = "nowrap")# %>% 


```


## Analysis
### Scaling

Values and weights are all measured relative to each other, so each participant must be using the same scale. Some respondents did not use 100 and 0 as the highest and lowest scores for all indicator value functions and weights. To make their relative values comparable to other indicators all value functions provided where scaled to 0-100. Likewise, weightings were scaled so that the highest indicator weighting for each panelist was 100.

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName"> Show/Hide details of responses requiring scalling </button>  
<div id="BlockName" class="collapse">  


**Value functions** scaled for:  

`r paste0("<br>- ", which.scaled.VF)`  

**Weights** were scaled for:  

`r paste0("<br>- ", which.scaled.wt)`  

### Displaying of results

Between rounds, expert-opinion estimates were presented on interactive plots to allow respondents to investigate estimates provided by others, and their certainties.

Value functions where plotted with lines connecting each of the points provided, and a model (details below) was fitted to present the average trend, where a consensus was deemed to be emerging. Hover-text displayed respondent name, their certainty in their estimate, and a sentence provided that describes the trend.

Indicator relative weightings were presented on a dot-plot, including the mean and median results for each indicator's weight.  

#### Model details

Average value functions were modeled with a binomial GAM. The influence of points from each respondent weighted by the inverse of the number of points provided by that respondent, with the aim of ensuring that all respondents had the same influence on the result, regardless of how many points they provided.


# Results

## Round 1
### Participants and survey completion

`r has_completed$respondent_names %>% unique() %>% length()` expert practitioners from Woodland Trust and Plantlife staff were invited to provide their opinion the value and weight of indicators.  
`r sum(completed_summary[,2:5] %>% rowSums() >0)` have provided a response as of `r date()`.   
Of these, `r sum((completed_summary[,2:5] %>% rowSums() >0) & (completed_summary[,2:5] %>% rowSums() < (4* has_completed$indicator_name %>% unique() %>% length() ) ))` remain partially incomplete.  
The chart below summarises participant completion of different elements of the survey. For more detail see table (button below chart).

``` {r  participation_chart, echo = F }

# use df to ggplot
# The first legitimate use of a pie chart in my carer!

max_count = max(completed_summary$vf)

create_gg_pie_chart <- function(count, max_count) {
  percentage <- count / max_count * 100
  pie_chart <- ggplot(data.frame(x = c("Done", "Remaining"), y = c(count, max_count - count)),
                      aes(x = "", y = y, fill = x)) +
    geom_bar(width = 1, stat = "identity") +
    scale_fill_manual(values = c("steelblue", "lightgray")) +  # Specify custom colors
    coord_polar(theta = "y") +
    theme_void() +
    theme(legend.position = "none")
  
  return(pie_chart)
  
}

gg_pie_charts <- list(NA)
bk = 1
for (i in 1:(length(completed_summary$respondent_names))){
  for (ii in 1:(dim(completed_summary)[2]-1)){
    count <- completed_summary[i,1+ii] %>% as.numeric()
    gg_pie_charts[[bk]] <-  create_gg_pie_chart(count, length(unique(df$indicator_name)))
  bk = bk+1
    }}
  

# Convert the list of pie charts to a grid using grid.arrange
pie_matrix <- arrangeGrob(grobs = gg_pie_charts, nrow = length(completed_summary$respondent_names), ncol = length(colnames(completed_summary[, c("wt", "wt_cert", "vf", "vf_cert")])))
# Add row names (respondent names)
row_names <- completed_summary$respondent_names

# Add column names
col_names <- c("Value func", "VF cert", "Weights", "wt cert")

combine <- rbind(tableGrob(t(c(col_names)), theme = ttheme_minimal(), rows = ""), 
                 cbind(tableGrob(row_names, theme = ttheme_minimal()), 
                       arrangeGrob(grobs = gg_pie_charts, nrow = length(row_names), ncol = length(col_names)),  size = "last"), size = "last")
grid.newpage()
grid.draw(combine)

```
Figure 1. Completion of expert opinion survey components by each panelist (coloured portion of pies) for value functions ("Value func"), value function certainties ("VF cert"), indicator weightings ("Weights") and weighting certainties ("wt certs").  

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName"> Show/Hide table of detailed completion info </button>  
<div id="BlockName" class="collapse">  

Details of incomplete sections

```{r what needs done by participants details}

# Create a data frame with participant names and the four columns
participant_table <- data.frame(respondent_names = completed_summary$respondent_names,
                                vf = "",
                                vf_cert = "",wt = "",
                                wt_cert = "",
 stringsAsFactors = FALSE)

# Iterate over each participant and update the table with the indicators not completed
for (i in 1:nrow(completed_summary)) {
  n.indicators <- has_completed$indicator_name %>% unique() %>% length()
  participant <- completed_summary[i, "respondent_names"] %>% as.character()
  not_completed_wt <- has_completed$sheet_name[has_completed$respondent_names == participant & has_completed$wt == FALSE]
  not_completed_wt_cert <- has_completed$sheet_name[has_completed$respondent_names == participant & has_completed$wt_cert == FALSE]
  not_completed_vf <- has_completed$sheet_name[has_completed$respondent_names == participant & has_completed$vf == FALSE]
  not_completed_vf_cert <- has_completed$sheet_name[has_completed$respondent_names == participant & has_completed$vf_cert == FALSE]
  
  if (length(not_completed_wt) == n.indicators) not_completed_wt <- "all"
  if (length(not_completed_wt_cert) == n.indicators) not_completed_wt_cert <- "all"
  if (length(not_completed_vf) == n.indicators) not_completed_vf <- "all"
  if (length(not_completed_vf_cert) == n.indicators) not_completed_vf_cert <- "all"

  participant_table[participant_table$respondent_names == participant, "wt"] <- paste(not_completed_wt, collapse = ", ")
  participant_table[participant_table$respondent_names == participant, "wt_cert"] <- paste(not_completed_wt_cert, collapse = ", ")
  participant_table[participant_table$respondent_names == participant, "vf"] <- paste(not_completed_vf, collapse = ", ")
  participant_table[participant_table$respondent_names == participant, "vf_cert"] <- paste(not_completed_vf_cert, collapse = ", ")
}

# Print the participant table

names(participant_table) <- c("Respondent", "Value Functions", "vf certs", "Weights", "wt certs")

participant_table %>%  
    datatable(extensions = c('FixedColumns',"FixedHeader"),
              options = list(scrollX = T, 
                             scrollY = '400px',  # Set the desired height for the table
                             scrollCollapse = TRUE,

                             paging = F,
                             fixedHeader=F,
                             dom = 't',
                             autoWidth = TRUE, fixedColumns = list(leftColumns = 1) #,
                             # fixedColumns = list(leftColumns = 1) ,
                             # buttons = c('copy', 'csv')
                             ),
              rownames = F#,
              #container = sketch.lcm.areas
              ) %>% 
    formatStyle( columns = c("Respondent"), fontWeight = 'bold', `text-align` = 'left', "white-space" = "nowrap")# %>% 

```

</div>
  <br/>
  

## Indicator value functions  

```{r make value function plots, echo = F }
gam_model <- list(NA)
dummy_data <- list(NA)
each.plotly <- list(NA)
each.plotly.no.model <- list(NA)
should.model <- list(NA)

for(i in 1: length(ind.matcher.df$indicator_name)){
  # Choose an indicator name
  indicator_name <- ind.matcher.df$indicator_name[i]
  # Filter the data for the selected indicator
  filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
    mutate(respondant_name = as.factor(respondant_name))
  # Color palette
  respondant_colours <- glasbey.colors(nlevels(as.factor(filtered_data$respondant_name))) %>% as.character()
  # respondant_colours <- brewer.pal(nlevels(as.factor(filtered_data$respondant_name)),  "Paired")

  # prediction of mean trend ----
  filtered_data$point.influence <- 1/(table(filtered_data$respondant_name)[as.factor(filtered_data$respondant_name)]) %>% 
    as.numeric()
  gam_model[[i]] <- gam(value.dec ~ s(measure, k = 4),
                        data = filtered_data, family = binomial(), weights = point.influence)
  # Create dummy data for prediction
  dummy_data[[i]] <- data.frame(measure = seq(min(filtered_data$measure), max(filtered_data$measure), length.out = 50))
  # Predict using the GAM model
  dummy_data[[i]]$predicted_value <- predict(gam_model[[i]], newdata = dummy_data[[i]], type = "response")*100
  
  # Plot of lines and mean predictions
  plot <- ggplot() +
    geom_line(data = dummy_data[[i]], size = 2, aes(y = predicted_value, x = measure), colour = "black") +
    geom_line(data = filtered_data, 
              aes(x = measure, y = value, color = respondant_name,
                  text = map(
                    paste0("<b>", `respondant_name`, "</b>",
                           "<br><b>Certainty:</b>", cert_val_funct, "<br>",
                           "<b>Weight: </b>", weight, " (certainty ", cert_weight,")<br>",
                           "<b>Sentance: </b>", vf.sentance),
                    HTML),
                  alpha = cert_val_funct),
              size = 0.5) +
    geom_point(data = filtered_data, 
               size = 2, shape = 16,
               aes(x = measure, y = value, color = respondant_name,
                   text = map(
                     paste0("<b>", `respondant_name`, "</b>",
                            "<br><b>Certainty:</b>", cert_val_funct, "<br>",
                            "<b>Weight: </b>", weight, " (certainty ", cert_weight,")<br>",
                           "<b>Sentance: </b>", vf.sentance),
                     HTML))) +
    scale_colour_manual(values = respondant_colours) +
    labs(title = indicator_name, x = ind.matcher.df$ind.axis.title[i], y = "Value Score",
         colour = "Respondant") +
    guides(alpha = F) +
    theme_pubr()+
    theme(plot.title = element_text(size = 12),
          plot.subtitle = element_text(size = 12),
          legend.title = element_text(size = 12 , face = "bold"),
          legend.text =  element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 12),
          axis.title =  element_text(size = 12),
          legend.position = "right",
          plot.margin = margin(0.5, 1, 0.5, 1, "cm")
    ) 
  each.plotly[[i]] <- ggplotly(plot, tooltip = "text", dynamicTicks = F) %>% 
    config(displayModeBar = F) %>% 
    layout(yaxis = list(range = c(-5, 105)))
  
  # Plot of lines and mean predictions
  plot <- ggplot() +
    # geom_line(data = dummy_data[[i]], size = 2, aes(y = predicted_value, x = measure), colour = "black") +
    geom_line(data = filtered_data, 
              aes(x = measure, y = value, color = respondant_name,
                  text = map(
                    paste0("<b>", `respondant_name`, "</b>",
                           "<br><b>Certainty:</b>", cert_val_funct, "<br>",
                           "<b>Weight: </b>", weight, " (certainty ", cert_weight,")<br>",
                           "<b>Sentance: </b>", vf.sentance),
                    HTML),
                  alpha = cert_val_funct),
              size = 0.5) +
    geom_point(data = filtered_data, 
               size = 2, shape = 16,
               aes(x = measure, y = value, color = respondant_name,
                   text = map(
                     paste0("<b>", `respondant_name`, "</b>",
                            "<br><b>Certainty:</b>", cert_val_funct, "<br>",
                            "<b>Weight: </b>", weight, " (certainty ", cert_weight,")<br>",
                           "<b>Sentance: </b>", vf.sentance),
                     HTML))) +
    scale_colour_manual(values = respondant_colours) +
    labs(title = indicator_name, x = ind.matcher.df$ind.axis.title[i], y = "Value Score",
         colour = "Respondant") +
    guides(alpha = F) +
    theme_pubr()+
    theme(plot.title = element_text(size = 12),
          plot.subtitle = element_text(size = 12),
          legend.title = element_text(size = 12 , face = "bold"),
          legend.text =  element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 12),
          axis.title =  element_text(size = 12),
          legend.position = "right",
          plot.margin = margin(0.5, 1, 0.5, 1, "cm")
    ) 
  each.plotly.no.model[[i]] <- ggplotly(plot, tooltip = "text", dynamicTicks = F) %>% 
    config(displayModeBar = F) %>% 
    layout(yaxis = list(range = c(-5, 105)))
  
  should.model[[i]] <- T

}

```



```{r vf plot vfs, echo = F , results="asis"}

for (i in 1:length(ind.matcher.df$indicator_name)) {
  cat("###", ind.matcher.df$sheet_name[i], "\n\n")
   print(htmltools::tagList(each.plotly[[i]] ))
   cat('\n')
}


```
```{r vf 2plot vfs, echo = F , results="asis"}

   # exceptions <- expert.data$value_func_exceptions[expert.data$indicator_num %>% unlist()==i] 
   # respondants <- expert.data$respondant_name[expert.data$indicator_num %>% unlist()==i] %>% unlist()
   # 
   # for ( ii in 1:sum(lengths(exceptions)>0) ){
   #   cat("*",respondants[lengths(exceptions)>0][ii], "*  ")
   #   exceptions[lengths(exceptions)>0][ii] %>% unlist()
   #   }

# }
# 
#  cat("  \n###",  month.name[i], "Air Quaility  \n")
#   #print(plot(airquality[airquality$Month == i,]))
#   plot(airquality[airquality$Month == i,])
#   cat("  \n")
# 
# 



# i <- 1
# 
# each.plotly[[i]]
# 
# exceptions <- expert.data$value_func_exceptions[expert.data$indicator_num %>% unlist()==i] 
# respondants <- expert.data$respondant_name[expert.data$indicator_num %>% unlist()==i] %>% unlist()
# 
# respondants[lengths(exceptions)>0]
# 
# for ( ii in 1:sum(lengths(exceptions)>0) ){
#   print(paste0("*",respondants[lengths(exceptions)>0][ii], "*  "))
#   exceptions[lengths(exceptions)>0][ii]
# 
# }



```

## Indicatior weights

```{r}
#-----------------------------------------------------
# PLOT WEIGHTS --------------------------------------------------
#-----------------------------------------------------

plot <- ggplot() +
  geom_boxplot(data = just.one.df, aes(x = sheet_name, y = weight), 
               width=0.3, color="grey70", fill = "white", alpha=0.2) +
  geom_point(data = just.one.df, 
             position = position_jitter(width = 0.2, height = 0),
             size = 1, shape = 16,
             aes(x = sheet_name, y = weight, color = respondant_name,
                 text = map(
                   paste0("<b>", `respondant_name`, "</b><br>",
                          "<b>Weight: </b>", weight, " (certainty ", cert_weight,")<br>"),
                   HTML))) +
  stat_summary(data = just.one.df, aes(x = sheet_name, y = weight,
                                       text = map(
                                         paste0("Mean weight"),
                                         HTML)),
               geom = "point", fun = mean,
               color = "red", shape = 3, size = 2,
               position = position_nudge(x = 0.05, y = 0)
               ) +
  labs(x = NULL, y = "Relative importance (weight)", 
       title = "Indicator weights",
       color = "Respondant") +
  theme_pubr()+
  theme(plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 12),
        legend.title = element_text(size = 12 , face = "bold"),
        legend.text =  element_text(size = 12),
        axis.text.x = element_text(angle = -60, hjust = 0, size = 10),
        axis.text.y = element_text(size = 12),
        axis.title =  element_text(size = 12),
        legend.position = "right",
        plot.margin = margin(0.5, 1, 0.5, 1, "cm")
  ) 
ggplotly(plot, tooltip = "text", dynamicTicks = F) %>% 
  config(displayModeBar = F) %>% 
  layout(yaxis = list(range = c(-5, 105)))
```


## Certainties



# Discussion points
- [ ] - Some value functions have a minimum score >0. Strictly speaking this is okay... but it does meant that the final condition score will not be on a 0-100 scale (there will always be some minimum condition score).








