# Canopy nattiness

## Justification

## Measuring

# Respondent value functions and weights

```{r load curated data, echo = F, include = F, eval = T }
extraction.location <- "Data\\Delphi round 1\\"
load(paste0(extraction.location,"curated.RData"))
```

```{r libraries_data, echo = F, include = F, eval = T }

# LIBARIES ---------------------------------------------------------------
# library(readxl)
library(tidyverse)
library(ggpubr)
library(plotly)
library(htmltools)

# library(stringr)
# library(plotly)
# library(htmlwidgets)
# library(gridExtra)
# library(grid)
# library(DT)
# library(ggiraph)
# library(cowplot)
# library(ggnewscale)
# library(Polychrome)
source("Scripts\\functions_delphi_analysis.R")

forms.direct <- "Data\\Delphi round 1\\response sheets\\"
extraction.location <- "Data\\Delphi round 1\\"

```

```{r vf Canopy Nativness - init, echo = F , results="asis"}

ind.num = 2
indicator_name = ind.matcher.df$indicator_name[ind.num]
ind.axis.title <- ind.matcher.df$ind.axis.title[ind.num]

filtered_data <- df[df$indicator_name == indicator_name, ] %>% 
  mutate(respondant_name = as.factor(respondant_name))

```

`r paste(length(unique(filtered_data$respondant_name)))` respondents.

::: panel-tabset
#### Continuous Value function

```{r vf - continuous, echo = F , results="asis"}
#| column: page

continuous_vf_fig()

```

#### Categorised value function

```{r vf - categorical, echo = F , results="asis"}
#| column: page

categorised_vf_fig()

## table
filtered_data %>%
  group_by(measure) %>%
  summarise(Mean = mean(value, na.rm = TRUE) %>% round(digits = 1),
            'Standard deviation' = sd(value, na.rm = TRUE) %>% round(digits = 1)) %>% 
knitr::kable()

```

#### Weights

```{r vf - weights, fig.width= 4, echo = F , results="asis"}
#| column: page

ggplot_resp_weight_ind(indicator_name = indicator_name)

```
:::

### Comments on data

# Respondent comments

### Considerations for indicator design

### Priority exceptions

### Lower priority exceptions

### Concerns already covered by this or other indicators

# Other approaches

## Woodland Wildlife Toolkit & NFI

## Current WT estate condition assessment

# References
